UBR-1 deficiency leads to ivermectin resistance in C. elegans

Version 1

Abstract

Abstract

Resistance to anthelmintics, particularly the macrocyclic lactone ivermectin (IVM), presents a substantial global challenge for parasite control. We found that the functional loss of an evolutionarily conserved E3 ubiquitin ligase, UBR-1, leads to IVM resistance in Caenorhabditis elegans. Multiple IVM-inhibiting activities, including viability, body size, pharyngeal pumping, and locomotion, were significantly ameliorated in various ubr-1 mutants. Interestingly, exogenous application of glutamate induces IVM resistance in wild-type animals. The sensitivity of all IVM-affected phenotypes of ubr-1 is restored by eliminating proteins associated with glutamate metabolism or signaling: GOT-1, a transaminase that converts aspartate to glutamate, and EAT-4, a vesicular glutamate transporter. We demonstrated that IVM-targeted GluCls (glutamate-gated chloride channels) are downregulated and that the IVM-mediated inhibition of serotonin-activated pharynx Ca2+ activity is diminished in ubr-1. Additionally, enhancing glutamate uptake in ubr-1 mutants through ceftriaxone completely restored their IVM sensitivity. Therefore, UBR-1 deficiency-mediated aberrant glutamate signaling leads to ivermectin resistance in C. elegans.

Introduction

Anthelmintic resistance (AR) to broad-spectrum antiparasitic macrocyclic lactones (MLs), which include ivermectin (IVM), avermectin (AVM), and doramectin (DOM), has emerged as a critical global concern in veterinary parasites (1–3). The rise of anthelmintic-resistant nematodes poses a pressing issue, impacting not only livestock production but also animal welfare and human health. Overcoming this challenge requires a comprehensive understanding of the molecular mechanisms driving nematode resistance, which, in turn, will facilitate the development of innovative and effective parasite prevention strategies.

The nematode Caenorhabditis elegans (C. elegans), owing to its sensitivity to a majority of anthelmintic drugs and conservation of functional genes, serves as a crucial model organism for identifying genes related to IVM resistance (4–7). Mutations in three C. elegans glutamate-gated chloride channel (GluCl) genes—avr-15, avr-14, and glc-1—have collectively demonstrated strong resistance to ivermectin (8–11). Exogenous expression and crystallization X-ray structural analysis support the premise that these GluCls act as the primary targets of IVM (12, 13). Further identification of IVM resistance-related genes, including P-glycoprotein transporters (pgp-1, pgp-3, pgp-6, pgp-9, pgp-13, and mrp-6), cytochrome oxidases (cyp-14, cyp-34/35), and glutathione S-transferases (gst-4, gst-10) (14–16), has revealed their role in enhancing ivermectin excretion or metabolism. These detoxification genes are thought to be regulated by the ubiquitous transcription factor NHR-8, which encodes a nuclear hormone receptor (17). While these studies have revealed a polygenic mechanism underlying IVM resistance, increasing reports of unknown causes of IVM resistance continue to emerge (18–20), suggesting that additional unknown mechanisms are awaiting investigation.

UBR1, an E3 ubiquitin ligase and a pivotal component of the ubiquitin‒ proteasome system, facilitates protein degradation (21–25). UBR1 is implicated in substrate-specific recognition and metabolism, primarily through the N-end rule (26). In humans, loss-of-function mutations in UBR1 result in Johanson-Blizzard syndrome (JBS), a rare autosomal recessive disorder characterized by a spectrum of developmental and neurological symptoms (27, 28). In the nematode Caenorhabditis elegans, a single UBR-1 homolog retains all major conserved functional domains (29). LIN-28 is the sole identified substrate of C. elegans UBR-1, acting as an RNA-binding pluripotency factor crucial for seam cell patterning (30). However, the absence of LIN-28 neither replicates nor ameliorates the movement defects observed in ubr-1 mutants-stiff body bending during backward movement (29).

Our recent research has consistently shown that UBR-1 regulates neural outputs through glutamate homeostasis mechanisms. Loss of UBR-1 leads to elevated glutamate levels in animals, disrupting coordinated motor patterns, likely due to a compensatory decrease in the expression of excitatory glutamate receptors (29). Furthermore, ubr-1 mutations disturb inherent asymmetric neural activity by activating inhibitory GluCls (GLC-3 and GLC-2/4), which in turn impairs rhythmic defecation motor programs (31). These observations suggest that aberrant glutamate metabolism resulting from UBR-1 deficiency has a widespread effect on glutamatergic signaling pathways. This leads to the question of whether ubr-1 mutants cause comprehensive changes in IVM-targeted GluCls, thereby contributing to IVM resistance.

To investigate this possibility, we assessed the IVM sensitivity of wild-type and ubr-1 mutant animals. We found that, in contrast to wild-type N2 animals, loss-of-function ubr-1 mutants present a range of IVM-resistant phenotypes. N2 animals exhibit high fatality rates, abnormal development and motor defects, such as pharyngeal pumping or locomotion disabilities, upon exposure to ivermectin. In contrast, ubr-1 mutants were almost insensitive to IVM. The elimination of the glutamate synthetase GOT-1 and the glutamate transporter EAT-4 completely reversed the IVM sensitivity of ubr-1 mutants, highlighting the critical role of the glutamatergic signaling pathway. Through the use of translational reporters and in situ calcium imaging, we determined that IVM resistance in ubr-1 is due to the functional downregulation of IVM-targeted GluCls, including AVR-15, AVR-14, and GLC-1. This downregulation is presumably also triggered by the aberrantly elevated glutamate-induced compensatory decrease, as the application of exogenous glutamate in wild-type animals partially replicated the IVM resistance observed in ubr-1 mutants. Moreover, pharmacologically reducing glutamate levels via the use of ceftriaxone (32, 33) successfully restored IVM sensitivity in ubr-1 mutants. Thus, our study presents multiple lines of evidence for a novel IVM resistance mechanism linked to the deficiency of the ubiquitin ligase UBR-1 via a glutamate metabolism pathway.

Results

Dose- and time-dependence of IVM resistance in ubr-1 mutants

To further elucidate the mechanisms of anthelmintic resistance (AR), we employed a simplified assay derived from prior research (Figure 1A). We assessed the survival of nematodes on media infused with antiparasitic macrocyclic lactones (MLs). In essence, synchronized L4-stage worms were placed on NGM plates with OP50 bacteria and treated with various MLs, including ivermectin (IVM, 5 ng/mL), avermectin (AVM, 5 ng/mL), or doramectin (DOM, 5 ng/mL). Twenty hours postexposure, we quantified the surviving worms and calculated the percentage of viability relative to the initial population.

ubr-1 exhibits IVM-resistant phenotypes.

(A) Left: Schematic representation of the IVM resistance test in C. elegans. Well-fed L4 stage animals were transferred to plates containing OP50 bacteria with or without macrocyclic lactones (MLs) for 20 hours. Right: Representative grid plots illustrating the viability of wild-type and ubr-1 animals in response to ivermectin (IVM, 5 ng/mL), avermectin (AVM, 5 ng/mL), and doramectin (DOM, 5 ng/mL). (B) Quantification analysis of the viability of wild-type and ubr-1(hp684) mutants exposed to different MLs. ubr-1(hp684) mutants exhibit resistance to various MLs. (C) Dose‒response curve depicting the viability of wild-type and ubr-1(hp684) mutants in the presence of varying concentrations of IVM. The IC50 was 5.7 ng/mL for the wild type and 11.6 ng/mL for the ubr-1 mutants. (D) Time-dependent effect of IVM exposure on animal viability. The IC50 values were 20.2 hours for the wild type and 42.1 hours for the ubr-1 mutants. (E) Representative images of worm size in the wild type and ubr-1(hp684) mutants with or without IVM treatment. Scale bar, 50 µm. (F) Quantitative analysis of body length in different genotypes with or without IVM. (G) Quantification of the average pharynx pump number in animals with or without IVM. (H) Raster plots illustrating the locomotion velocity of individual animals in the absence (0 ng/mL) and presence (5 ng/mL) of IVM. n = 10 animals in each group. (I) Quantification of the average velocity in different genotypes with or without IVM treatment. For the viability test, n = 50 animals per plate, which was repeated at least 5 times (5 trials). ns, not significant, *p < 0.05, **p < 0.01, ***p < 0.001 by Student’s t test. The error bars represent the SEM.

Our observations revealed a significant decline in the viability of wild-type nematodes following ML exposure (IVM 52.0 ± 3.2%, AVM 51.6 ± 2.9%, DOM 52.4 ± 3.4%, in contrast to the vehicle control 100 ± 0.0%), confirming the susceptibility of C. elegans to MLs (Figure 1B). In contrast, ubr-1(hp684; lf) mutants presented remarkable resistance to ML treatments (IVM, ubr-1 96.0 ± 1.7% vs vehicle control 100 ± 0.0%) when juxtaposed with their wild-type counterparts (IVM, WT 52.0 ± 3.2%) (Figure 1B). This trend of increased survival was consistently observed with ubr-1(hp684) mutants treated with AVM (WT 51.6 ± 2.9% vs ubr-1 82.8 ± 2.9%) and DOM (WT 52.4 ± 3.4% vs ubr-1 83.6 ± 3.1%). These findings reveal the pronounced ML resistance characteristics of ubr-1 loss-of-function mutants.

We further investigated the dose- and time-dependent resistance of ubr-1(hp684) to select IVM as a proxy for MLs because of its efficiency and common application. We tested a spectrum of IVM concentrations ranging from 1 ng/mL to 11 ng/mL (Figure S1A). The data revealed that ubr-1(hp684) mutants exhibited resistance at all tested doses (Figure 1C), resulting in a doubling of the half-maximal inhibitory concentration (IC50) — from 5.7 ng/mL in wild-type N2 worms to 11.6 ng/mL in ubr-1 mutants. Time-dependent IVM resistance was also compared between the wild type and ubr-1 mutants (Figure S1B). The duration of half-maximal inhibition in the wild type was approximately 20.2 hours, which increased to 42.1 hours in the ubr-1(hp684) mutants (Figure 1D). To highlight the disparity in IVM resistance, we utilized an assay with 5 ng/mL IVM for a 20-hour period (indicated by the vertical dashed lines in Figure 1C and D) for subsequent experiments, unless otherwise specified.

Additional IVM resistance phenotypes of ubr-1

In addition to viability, ubr-1 mutants maintained normal growth despite IVM exposure. After 20 hours of IVM, the body size of the wild-type animals significantly decreased (34), resulting in a marked decrease in the middle line body length (control, 1.07 ± 0.02 mm; IVM, 0.62 ± 0.03 mm) (Figure 1E). In contrast, the size of the ubr-1 mutants (control, 1.08 ± 0.02 mm; IVM, 1.00 ± 0.02 mm) was largely preserved, indicating that the ubr-1 mutants developed into typical young adults (Figure 1F). Notably, while IVM exposure nearly halted pharyngeal pumping in wild-type animals (control 247.4 ± 4.1 min-1, IVM 34.6 ± 2.8 min-1), ubr-1 mutants sustained considerable pumping activity (control 254.1 ± 5.0 min-1, IVM 152.9 ± 3.0 min-1) (Figure 1G). Motor functions were similarly resistant to IVM in ubr-1 mutants (Movie S1). The velocity of free movement under IVM treatment was significantly greater in ubr-1 mutants (135.9 ± 7.2 μm/s) than in wild-type animals (23.8 ± 2.1 μm/s) (Figure 1H, I, Figure S1C). No significant differences in viability, body length, pharyngeal pumping, or locomotion speed were detected between ubr-1 and wild-type animals in the absence of IVM treatment.

These results collectively indicate that ubr-1 loss-of-function mutants present a comprehensive spectrum of phenotypes associated with IVM resistance.

Different ubr-1 mutants exhibit consistent IVM resistance

The ubr-1 gene is functionally conserved from yeast to humans. In nematodes or filarial worms, ubr-1 also exhibited high sequence similarity (Figure S2A). The hp684 allele contains a premature stop codon, resulting in the truncation of the final 194 amino acids of UBR-1 (Q1864X) (Figure S2B). To validate functional congruence, we expanded our study to include additional ubr-1 mutant alleles. These alleles, which are genetically predicted to lack one or more critical conserved domains, including hp865, which replaces the entire RING finger domain with an SL2-NLS::GFP, and hp821 hp833 (E34X, E1315X), each introducing stop codons that suggest the absence of all functional domains. All these alleles demonstrated IVM resistance comparable to that of the hp684 allele. Specifically, the viability of hp865 (81.6 ± 3.8%) and hp821 hp833 (76.4 ± 2.6%) increased following IVM treatment (Figure S2C). Similarly, the body lengths of hp865 (0.93 ± 0.02 mm) and hp821 hp833 (0.90 ± 0.02 mm) were not diminished by IVM exposure (Figure S2D). Pharyngeal pumping and spontaneous movement velocities were also significantly greater in hp865 (154.2 ± 3.4 min-1, 100.2 ± 7.4 μm/s) and hp821 hp833 (134.9 ± 3.7 min-1, 86.1 ± 6.9 μm/s) than in the wild type (Figure S2E, F). These findings demonstrate that various ubr-1 mutants exhibit a consistent IVM-resistant phenotype.

Glutamate induces IVM resistance in wild-type worms

Systemic elevation of glutamate levels in ubr-1 mutants has been documented (29, 35, 36), prompting an investigation into whether this increase is the key determinant of the observed resistance to IVM.

To test this hypothesis, we pretreated wild-type N2 animals with glutamate to mimic the excessive glutamate environment in ubr-1 mutants (Figure 2A). During these trials, wild-type larvae were exposed to L-glutamate (20 mM) on NGM plates from the L1 to L4 stages before their transfer to IVM-containing plates. Strikingly, these glutamate-pretreated animals presented a significant increase in IVM resistance (Figure 2B).

Glutamate mimics IVM resistance in wild-type N2 animals, and ceftriaxone fully reverses IVM resistance in ubr-1 mutants.

(A) Modified schematic representation of the pretreatment approach in the IVM resistance test in C. elegans. Synchronized L1 animals were cultured on plates containing OP50 bacteria and pretreated with glutamate (Glu, 5 mM), γ-aminobutyric acid (GABA, 5 mM), or aspartate (Asp, 5 mM) until they reached the L4 stage. Pretreated animals were then transferred to plates containing additional IVM (5 ng/mL) for 20 hours. (B) Representative grid plots illustrating the viability of wild-type and ubr-1 animals pretreated with different glutamate metabolites. (C-F) Quantitative analysis of viability, body length, pharyngeal pump rate, and locomotion velocity in wild-type and ubr-1 mutants following exposure to various glutamate metabolites. Wild-type worms treated with glutamate displayed resistance to IVM. However, glutamate did not affect the IVM resistance of the ubr-1 mutant. ns, not significant, ****p < 0.0001 by one-way ANOVA in C-F. (G) Diagram illustrating the ceftriaxone pretreatment procedure in the IVM resistance test. Synchronized L1 animals were cultured on OP50-fed plates with or without ceftriaxone (50 μg/mL) until the L4 stage. The animals were subsequently transferred to plates seeded with additional IVM (5 ng/mL) for 20 hours. (H) Upper: Representative grid plots depicting the viability of ubr-1 animals at different concentrations of ceftriaxone. Bottom: Dose‒response curve illustrating the amount of ceftriaxone required for viability in ubr-1(hp684) mutants (IC50 = 5.1 μg/mL). (I-L) Quantification of viability, body length, pharyngeal pumping, and locomotion velocity in the wild type and ubr-1 mutants in the absence and presence of ceftriaxone. Ceftriaxone fully restored the sensitivity of ubr-1 mutants to IVM, comparable to the levels observed in wild-type N2 animals. ns, not significant, ***p < 0.001 by Student’s t test in I-L. Error bars represent SEM.

Specifically, the viability of these glutamate-exposed animals increased from 52 ± 3.2% to 77.2 ± 1.5% compared with that of their nonexposed counterparts (Figure 2C). A corresponding increase in body size was noted, from 0.62 ± 0.03 mm to 0.76 ± 0.01 mm (Figure 2D). Motor activity similarly improved, with velocities increasing from 23.79 ± 2.1 μm/s to 62.4 ± 7.6 μm/s (Figure 2E-F). These findings clearly demonstrate that the external application of glutamate can induce IVM resistance in wild-type N2 animals. Interestingly, in ubr-1 mutants, pretreatment with glutamate did not further increase IVM resistance, implying potential saturation of the glutamate effect in ubr-1 mutants (Figure 2B-F).

In addition to glutamate, other metabolically related molecules, such as aspartate (Asp), a precursor in glutamate synthesis, and γ-aminobutyric acid (GABA), a product of glutamate metabolism, were also found to be elevated in the ubr-1 mutant (29, 31). However, the cocultivation of wild-type N2 worms with aspartate or GABA did not confer IVM resistance (Figure 2A-F), suggesting the glutamate-specific regulation of IVM resistance. This finding highlights the specificity and sufficiency of glutamate in conferring IVM resistance in a wild-type context.

Enhancing glutamate uptake restores IVM sensitivity in ubr-1

Given the likelihood that elevated glutamate levels underpin IVM resistance in ubr-1, mitigating this glutamate surplus could reverse resistance and reinstate IVM sensitivity. To evaluate this premise, we employed ceftriaxone (Cef) (Figure 2G), which is known for enhancing glutamate uptake by upregulating excitatory amino acid transporter-2 (EAAT2) (32, 33), with the aim of reducing glutamate levels.

The results were striking: the IVM resistance typically observed in ubr-1 mutants was completely counteracted by Cef treatment. While Cef (50 μg/mL) had a negligible effect on wild-type N2 animals, it fully restored IVM sensitivity in ubr-1 mutants (Figure 2H). Moreover, this restoration followed a dose-dependent pattern, with a half-effect concentration of 5.14 μg/mL (Figure 2H). Cef treatment resulted in a decrease in the viability of ubr-1 mutants from 96.0 ± 1.7% to 62.8 ± 2.3% in the presence of IVM (Figure 2I). Additional IVM resistance phenotypes in ubr-1 mutants, such as body size, pharyngeal pumping frequency, and locomotion velocity, were fully ameliorated by Cef (Figure 2J-L).

In line with its classification as a beta-lactam antibiotic, Cef effectively reduced the growth of the worm food source E. coli OP50, resulting in a reduced bacterial lawn on NGM plates (Figure S3A, B). However, simply diminishing the OP50 bacterial quantity on NGM plates without Cef did not restore IVM sensitivity in ubr-1 mutants (Figure S3C-F), suggesting that food scarcity is not a critical factor. Similarly, increasing the number of OP50 bacteria present on NGM plates containing Cef did not affect its ability to restore IVM sensitivity in ubr-1 mutants (Figure S3C-F). These findings indicate that the antimicrobial action of ceftriaxone is not related to its ability to restore IVM sensitivity in ubr-1 mutants.

Hence, these results indicate that pharmacologically reducing glutamate levels in ubr-1 reestablishes IVM sensitivity, leading to the hypothesis that aberrant glutamate metabolism underlies IVM resistance in ubr-1.

Restoration of ubr-1 IVM sensitivity through glutamatergic signaling pathway regulation

Our previous research revealed the involvement of UBR-1 in regulating glutamate metabolism, which is essential for coordinated reversal bending and rhythmic defecation motor programs (29). The absence of UBR-1 also leads to perturbations in the balance between the GABAergic and glutamatergic signaling pathways (31). In C. elegans, IVM sensitivity is marked by the sustained activation of glutamate-gated chloride channels (GluCls) by IVM (11). These glutamate-related findings led us to explore whether the resistance of ubr-1 mutants to IVM could be attributed to alterations in the glutamate signaling pathway.

At glutamatergic synapses, a series of enzymatic reactions link glutamate to neurotransmitter recycling. In C. elegans, glutamate-oxaloacetate transaminases, or GOT enzymes, facilitate amino group conversion between aspartate and α-ketoglutarate (α-KG) to oxaloacetate (OAA) and glutamate (Figure 3A) (37, 38). We began our exploration by assessing IVM sensitivity following the disruption of GOT-1. While single got-1 loss-of-function mutants displayed IVM sensitivity similar to that of the wild type, ubr-1; got-1 double mutants presented a significant increase in IVM sensitivity, reverting to wild-type levels of viability, body length, and motor activity, in stark contrast to ubr-1 single mutants (Figure 3B-E). These findings indicate that impeding glutamate synthesis can effectively restore IVM sensitivity in ubr-1 mutants.

Deletion of got-1 and eat-4 suppresses ubr-1’s IVM resistance.

(A) Schematic of the signaling pathway for glutamate synthesis through transaminase GOT-1 and loading via vesicular transport VGLUT/EAT-4. (B-E) Quantification of viability, body length, pharyngeal pump rate, and locomotion velocity in different genotypes. The removal of got-1 and eat-4 significantly restored sensitivity to IVM in ubr-1 mutants. (F-I) IVM resistance phenotypes were rescued by expression of UBR-1 (+ UBR-1) driven by its own promoter (Pubr-1) or the pharynx muscle-specific promoter (Pmyo-2). For the viability test in F, n = 50 animals per plate and repeated at least 5 times (5 trials), n ≥ 10 animals in G-I, *p < 0.05, **p < 0.01, ***p < 0.001 by one-way ANOVA. Error bars, SEM.

Glutamate, a crucial excitatory neurotransmitter, is packed into synaptic vesicles by the vesicular glutamate transporter (VGLUT) and released into the synaptic cleft to activate postsynaptic receptors. We next explored the impact of inhibiting glutamate transport in ubr-1 mutants by eliminating EAT-4, the sole VGLUT in C. elegans (39).

Notably, the absence of EAT-4 in ubr-1 mutants dramatically reversed their IVM resistance, as evidenced by increased viability (ubr-1; eat-4 66.8 ± 3.5%), reduced body length (ubr-1; eat-4 0.56 ± 0.02 mm), increased pharyngeal pumping (ubr-1; eat-4 46.7 ± 3.6 min-1), and improved locomotion speed (ubr-1; eat-4 83.5 ± 7.4 μm/s) in ubr-1; eat-4 double mutants (Figure 3B-E).

In summary, these results underscore the importance of glutamate synthesis and synaptic transport in restoring IVM sensitivity in ubr-1 mutants, suggesting a role for synaptic glutamate signaling in the resistance of ubr-1 to IVM. The IVM resistance phenotypes of wild-type N2 animals were not rescued by ceftriaxone (Figure 2I-L), suggesting that ceftriaxone appears to be effective only against excess glutamate. This discovery is particularly intriguing, as it reveals a previously unexplored role of glutamate metabolism in the regulation of IVM resistance.

UBR-1 appears to regulate IVM resistance from the pharynx

UBR-1 is ubiquitously expressed across a variety of tissues, including the pharynx, body wall muscles, and neurons (Figure S4A) (29, 35, 36). To pinpoint the critical tissue for the role of UBR-1 in IVM resistance, we conducted tissue-specific rescue assays via a functional plasmid. Our findings revealed that restoring UBR-1 expression under its endogenous promoter attenuated the IVM resistance observed in ubr-1 mutants. Notably, when UBR-1 expression was driven by an exogenous pharyngeal muscle-specific promoter, it fully corrected the IVM resistance phenotypes of the ubr-1 mutants (Figure 3F-I). These results imply that although UBR-1 is important in multiple tissues, its role in the pharynx is particularly critical in modulating IVM sensitivity.

To verify the tissue-specific function of UBR-1, we utilized RNA interference (RNAi) to selectively decrease ubr-1 expression in the pharynx. This precise RNAi application to the pharynx indeed induced IVM resistance in wild-type N2 worms (Figure S4B-E). RNAi-mediated knockdown of ubr-1 in nonpharyngeal tissues, such as neurons and body wall muscles, also elicits an IVM resistance profile. Our research revealed that animals in which ubr-1 was knocked down specifically in the pharynx presented more pronounced IVM resistance phenotypes than did those in which ubr-1 was knocked down in other tissues. This phenotype approached the level observed when ubr-1 was knocked down via its native promoter, underscoring the unique contribution of the pharynx to ubr-1-mediated IVM resistance.

Taken together, these results confirm that UBR-1 predominantly regulates IVM resistance through its action in the pharynx.

Downregulation of IVM-targeted GluCls in ubr-1

What mechanisms underlie the induction of IVM resistance by abnormal glutamate metabolism? Previously, we established that ubr-1 mutants exhibit impaired GLR-1 glutamate receptor function in AVA/AVE interneurons (29). Three GluCls (avr-15, avr-14, glc-1) are recognized as pivotal for IVM resistance in C. elegans (11). We sought to determine whether compromised IVM-targeted GluCls contribute to IVM resistance in ubr-1 mutants.

To test this hypothesis, we created three transgenic strains with translational reporters for each GluCl under the control of their native promoters (Pavr-15::AVR-15::GFP, Pavr-14::AVR-14::GFP, Pglc-1::GLC-1::GFP). Notably, AVR-15::GFP expression was pronounced in the pharynx, whereas GLC-1::GFP displayed a distinct head expression pattern, particularly concentrated in the head body wall muscles (Figure 4A, Figure S5A). AVR-14::GFP expression was localized primarily to various head and body neurons (Figure S5A). Intriguingly, compared with that in wild-type animals, the fluorescence intensity of these GluCls was markedly lower in ubr-1 mutants (Figure S5A, B). The modulation of GluCl expression by ubr-1 seems to be posttranscriptional, given that the RNA levels of the GluCl-encoding genes (avr-15, avr-14, glc-1) did not significantly differ (Figure S5C).

Downregulation of AVR-15 in ubr-1 mutants.

(A) Representative fluorescence intensity of AVR-15 in different genotypes. (B) Quantitative analysis of normalized fluorescence intensity revealed a reduction in AVR-15::GFP expression in ubr-1 mutants. The diminished AVR-15 levels were rescued by overexpression of UBR-1 and genetically suppressed in got-1 and eat-4 mutants. Pharmacological intervention with ceftriaxone (50 μg/mL) successfully reinstated the pharyngeal expression of AVR-15::GFP. Conversely, glutamate administration reduced AVR-15 expression in the wild-type N2 strain. ROI, region of interest; n ≥ 10 animals in each group. **p < 0.01, ****p < 0.0001 by one-way ANOVA. Error bars, SEM.

Crucially, the reduction in pharyngeal AVR-15::GFP fluorescence observed in ubr-1 mutants was reversed upon reintroduction of UBR-1 (Figure 4A). Additionally, eliminating got-1 and eat-4 in these mutants mitigated the decrease in AVR-15::GFP expression within the pharynx (Figure 4A, B), suggesting a direct role of glutamate signaling in regulating GluCl expression levels. Notably, ceftriaxone pretreatment effectively reversed the diminished pharyngeal AVR-15::GFP levels in ubr-1 mutants (Figure 4A, B. In addition to glutamate pretreatment eliciting IVM resistance in wild-type animals, the application of glutamate markedly diminished the pharyngeal AVR-15::GFP fluorescence in these same wild-type worms (Figure 4A, B).

To further investigate the potential regulatory influence of ubr-1 on a broader spectrum of synaptic receptors, we selected two distinct categories of receptors that are not activated by glutamate. This reduction was not observed for the excitatory acetylcholine receptor (UNC-29/AChR) or inhibitory GABA receptor (UNC-49/GABAAR) (Figure S6), suggesting that ubr-1 may selectively affect glutamate-gated receptors (29).

Collectively, these results demonstrate that aberrant glutamate metabolism in the ubr-1 mutant leads to the downregulation of IVM-targeted GluCls, which likely contributes to diminished IVM responsiveness, culminating in the IVM resistance observed in ubr-1 mutants.

IVM completely suppresses serotonin-evoked pharynx activity in wild-type animals but not in ubr-1

IVM acts by engaging target GluCls, leading to the irreversible inhibition of pharyngeal pumping, disrupting conduction between nerve and muscle cells and eventually killing parasites (8, 11). The intricate dynamics of pharyngeal pumping in C. elegans rely on the coordinated interaction between the pharyngeal muscles and the excitatory pacemaker neuron MC, which is activated by serotonin (40).

To further examine the functional impairments of IVM-targeted GluCls, we conducted a direct examination of the impact of IVM on pharyngeal muscle cell activity. We generated a transgenic line expressing a Ca2+ sensor, GCaMP6, in the pharynx, enabling real-time monitoring of pharyngeal muscle activity (Figure 5A, Methods). Using 5-HT to stimulate dissected MC neurons, we induced intrinsic Ca2+ activity within the pharyngeal muscles. In these assays, 5-HT triggered strong Ca2+ transients across the pharyngeal muscle cells, from the terminal bulb through the isthmus to the anterior bulb (Figure 5A). To align with our quantified pumping behavior analysis, we focused our observations on the Ca2+ intensity within the terminal bulb. We noted a dose-dependent response of pharyngeal muscle cells to 5-HT, with an estimated EC50 of approximately 20.1 mM (Figure 5B, C). This assay provides a practical method to evaluate the functional impairment of IVM-targeted GluCls in ubr-1.

Reduced inhibition of serotonin-stimulated pharyngeal activity by IVM in ubr-1 mutants.

(A) Left: Schematic diagram of the pharynx motor circuit composed of an excitatory cholinergic motor neuron MC, an inhibitory glutamatergic motor neuron M3 and a group of compact muscles. Both types of motor neurons are innervated by serotonin (5-HT). Right: Representative pharynx muscle Ca2+ response and kymograph evoked by 5-HT (20 mM). The dashed lines aligned with the arrows denote the initiation of perfusion. (B) Average Ca2+ traces evoked by different concentrations of 5-HT. (C) Dose‒response curve of 5-HT for the pharynx muscle Ca2+ response in wild-type animals (EC50 = 20.1 mM). n ≥ 5 animals in each group. (D) Baseline-subtracted kymograph of the 5-HT-evoked pharynx muscle Ca2+ response in wild-type and ubr-1 mutants. IVM (3 μM) completely inhibited the wild-type pharynx Ca2+ response, which was significantly reduced in ubr-1 mutants. (E) Average Ca2+ traces evoked by 5-HT in different genotypes with or without IVM. (F) The dose-dependent inhibitory effect of IVM on pharynx Ca2+ was attenuated in ubr-1 mutants (blue line) compared with wild-type animals (black line) (IC50 = 0.5 μM in WT, while IC50 = 10.1 μM in ubr-1). (G) The decreased IVM inhibition of serotonin-evoked pharynx activity in ubr-1 was rescued by the restoration of UBR-1 expression (+ UBR-1), genetic removal of got-1 and eat-4, and pretreatment with ceftriaxone (+ Cef, 50 μg/mL). (H) IVM-mediated inhibition of serotonin-evoked pharynx activity in wild-type animals was blocked by glutamate feeding (WT + Glu, 20 mM). n = 10 animals in each group. (I) Quantification of the 5-HT-evoked peak amplitude of the pharynx muscle Ca2+ response in the different genotypes. Black stars: ****p < 0.0001 was used to compare intragroup differences or pharmacological groups via Student’s t test; apricot stars: ****p < 0.0001 was used to compare the differences with ubr-1 in the “+ 5-HT & IVM” group via one-way ANOVA. Error bars, SEM.

Interestingly, we observed that 5-HT induced similar Ca2+ transients in both wild-type and ubr-1 mutant worms. Furthermore, upon application of IVM (3 μM), 5-HT-elicited Ca2+ activity was profoundly suppressed in the wild-type pharynx (Figure 5D), mirroring its effect on pharyngeal pumping. However, ubr-1 mutants exhibited marked resistance to the suppressive effects of IVM, with only a modest decrease in the amplitude of Ca2+ transients (Figure 5E). Essentially, the capacity of IVM to inhibit 5-HT-driven pharynx activity was significantly compromised in ubr-1 mutants. Additionally, consistent with its behavioral impact, IVM dose-dependently suppressed 5-HT-elicited pharynx Ca2+ activity, with an estimated IVM IC50 of approximately 0.5 μM for wild-type and 10.1 μM for ubr-1 mutants (Figure 5F). These findings reveal the functional deficits of IVM-targeted GluCls in ubr-1.

In line with our earlier behavioral findings, the restoration of UBR-1 expression successfully reversed the inhibition of 5-HT-induced pharynx Ca2+ by IVM in ubr-1 mutants (Figure 5G). This corrective effect was similarly observed in ubr-1; got-1 and ubr-1; eat-4 double mutants (Figure 5G, I), reinforcing the association between decreased GluCl expression and attenuated IVM inhibition in ubr-1 mutants. These results suggest a mechanism whereby elevated glutamate levels in ubr-1 mutants may lead to glutamate-induced downregulation of GluCls, ultimately leading to IVM resistance.

To substantiate this proposed mechanism, we evaluated pharyngeal Ca2+ activity in ubr-1 mutants pretreated with Cef and in wild-type animals pretreated with glutamate. Consistent with our behavioral data, Cef treatment effectively reversed the inhibitory effect of IVM on 5-HT-evoked Ca2+ activity in the pharynx of ubr-1 mutants (Figure 5G, I), whereas glutamate exposure resulted in significant IVM resistance in wild-type animals (Figure 5H, I). This alignment between behavioral outcomes and Ca2+ activity measurements further supports the notion that glutamate accumulation is a critical factor in the downregulation of GluCls and the ensuing development of IVM resistance in ubr-1 mutants.

Discussion

In this study, we identified a crucial role of the E3 ubiquitin ligase UBR-1 in C. elegans, which contributes to IVM resistance. Given the importance of UBR-1 in regulating glutamate metabolism, our research introduces a novel mechanism for IVM resistance that could be applicable to other macrocyclic lactone (ML) anthelmintic resistance scenarios. Notably, this glutamate-based mechanism operates upstream, affecting the expression levels of GluCls, a departure from the previously documented variations in detoxification transporters that facilitate drug efflux. Through detailed investigation, we demonstrated that UBR-1 is instrumental in maintaining physiological glutamate homeostasis, thereby governing the abundance of various glutamate receptors. In ubr-1 mutants, perturbations in glutamate metabolism lead to reduced expression of primary IVM-sensitive GluCls, leading to decreased IVM-induced inhibition of serotonin (5-HT)-evoked pharynx Ca2+ activity and subsequent resistance to IVM.

A working model for UBR-1-mediated IVM resistance in C. elegans

Building upon our findings, we propose a working model that illustrates the role of the E3 ubiquitin ligase UBR-1 in mediating IVM resistance (Figure 6). In C. elegans, UBR-1 ensures functional glutamate homeostasis by inhibiting the activity of transaminase GOT-1 through a signaling pathway that has yet to be resolved. A deficiency in UBR-1 leads to uncontrolled production of excess glutamate by GOT-1, which is then packed into synaptic vesicles for release via the vesicular glutamate transporter VGLUT/EAT-4. This excess glutamate results in the downregulation of glutamate receptors, including the IVM-targeted GluCls AVR-15, AVR-14, and GLC-1. The consequent reduction in these receptors diminishes the inhibitory effect of IVM on 5-HT-evoked pharynx Ca2+ activity, highlighting the notable IVM resistance in ubr-1 mutants. Furthermore, the exogenous application of glutamate induces IVM resistance in wild-type animals, whereas the pharmacological enhancement of glutamate transporter activity via ceftriaxone mitigates resistance in ubr-1 mutants. These observations support the conclusion that IVM resistance in ubr-1 mutants is intricately linked with aberrant glutamate metabolism.

Unveiling a novel mechanism of anthelmintic resistance: aberrant glutamate metabolism

Glutamate is a central metabolic hub that connects glucose and amino acid metabolism in neurons and astrocytes. Disruptions in glutamate metabolism can precipitate an overabundance of glutamate, which in turn may trigger activity-dependent modulation of glutamate receptors. This phenomenon has been observed in Drosophila, where an increase in presynaptic glutamate concentrations leads to a notable reduction in both the number of glutamate receptors and the size of the synaptic field (41). Similarly, glutamate treatments have been shown to cause a loss of postsynaptic glutamate receptors (42). In mammalian systems, prolonged exposure to glutamate has been demonstrated to specifically reduce RNA editing levels of AMPA receptors in primary cortical neurons (43). These findings suggest that a compensatory downregulation mechanism may be responsible for the observed decrease in glutamate receptors in response to excess glutamate.

In our study, we revealed that an overproduction of glutamate resulting from ubr-1 mutations leads to the downregulation of IVM-targeted GluCls, which is subsequently linked to IVM resistance in C. elegans. This contrasts with previously elucidated mechanisms of IVM resistance, such as alterations in IVM-receptor binding sites that diminish their affinity, null mutations in IVM-targeted receptors, the upregulation of cellular efflux pumps, and enhanced drug degradation (2). Our research introduces a novel metabolic mechanism involving glutamate receptors. We demonstrated that the levels of IVM-targeted receptors are subject to modulation via the glutamate metabolic pathway. This finding highlights the potential for glutamate accumulation, whether through metabolic dysregulation or exogenous administration, to confer IVM resistance, thereby offering fresh perspectives on the mechanisms behind anthelmintic resistance. Our findings therefore reveal a new UBR-1-mediated pathway via GluCls that connects glutamate metabolism with IVM resistance.

Given the high degree of conservation in glutamate signaling pathways between C. elegans and parasitic nematodes, the implications of our findings are far-reaching. This highlights the importance of developing new pharmacological agents aimed at modulating glutamate metabolism for the control of parasitic nematodes. Moreover, altering environmental glutamate concentrations—whether through modifications in the physiological environment of nematodes or changes in the host’s diet—may increase the effectiveness of traditional macrocyclic lactone anthelmintics.

Expanding parasitic control through glutamate clearance

The recommendation for combination anthelmintic treatments to slow the emergence of anthelmintic resistance is well established, as they cover a similar range of parasites but act via different mechanisms (2, 44). Our research proposes an innovative approach for parasite prevention: the clearance of glutamate. The use of ceftriaxone, which has been shown to completely recover IVM sensitivity in ubr-1 mutants, presents a viable option to mitigate anthelmintic resistance prompted by excess glutamate. A combination treatment that includes ceftriaxone and IVM-like anthelmintics is advisable. Moreover, the concurrent use of ceftriaxone with other antiparasitic drugs may prove effective in treating AR parasites.

This dual-treatment strategy offers several benefits, such as the widespread availability of ceftriaxone, its established FDA approval since 1982 (45–47), the relatively low dose needed to increase IVM sensitivity, and its cost-effectiveness, particularly in the developing world.

Although the dynamics of anthelmintic resistance may vary among different hosts, our study contributes a fresh outlook by revealing a new mechanism of IVM resistance and advocating alternative preventative measures. We hope that these insights will attract the attention of global health authorities such as the World Health Organization and spark additional research in the realm of parasite control.

Materials and methods

C. elegans Strains and Transgenic Lines

C. elegans strains were cultured at 22°C on nematode growth medium (NGM) plates seeded with E. coli OP50 as a food source (48). The wild-type Bristol N2 strain was used, and the ubr-1(hp684) strains were sourced from an EMS mutagenesis screen. Additional ubr-1 alleles were generated via the CRISPR-Cas9 system (29). Other genetic mutants were obtained from the Caenorhabditis Genetics Center (CGC). Transgenic worms were generated through microinjection via standard protocols, with target DNA plasmids at ∼50 ng/µL coinjected with the marker plasmid Pmyo-2::RFP at 5–10 ng/µL. Table S1 shows a detailed list of the strains used in this study.

Ivermectin sensitivity assay

Ivermectin and avermectin (product number: PHR1380/31372) were purchased from Sigma Aldrich (St. Louis, MO, USA) and then diluted in dimethylsulfoxide (DMSO). Doramectin (product number: D127678) was purchased from Aladdin (Shanghai, China). A stock solution of 40 mM IVM was prepared in 1 mL of DMSO. Initial ML resistance assays utilized IVM, avermectin, and doramectin at 5 ng/mL. Standard NGM plates with OP50 and 0.03% DMSO were used, with varying IVM concentrations for tests. L4 hermaphrodites were synchronized and exposed to IVM for viability assays. After 20 hours, the body length, pharyngeal pump rate, and locomotion velocity of the survivors were analyzed via a stereomicroscope (Axio Zoom V16, Zeiss). Each experiment was independently replicated at least three times.

Animal viability and motor analysis

Animal viability was assessed by calculating the relative residual ratio: the number of viable worms after 20 hours of IVM incubation divided by the initial worm count. Viability (%), body length (mm), and velocity (µm/s) were measured via a custom MATLAB script (49).

For movement analysis, IVM-exposed worms were transferred to fresh OP50 plates. For locomotion velocity analysis, a single hermaphrodite was transferred to a 60 mm imaging plate. One minute after the transfer, a three-minute video of the crawled animal was recorded on a modified stereomicroscope (Axio Zoom V16, Zeiss) with a digital camera (acA2500-60um, Basler). Postimaging analyses utilized an in-house written MATLAB script. The central line was used for tracking. Images for velocity analysis from each animal were divided into 33 body segments. The midpoint was used to calculate the body length and velocity between each frame. The pharyngeal pumping frequency was directly calculated manually via video as pumps per minute (min-1).

Pharmacological pretreatment

For glutamate application, glutamate (product number: G1251, Sigma Aldrich) was dissolved in NGM at final concentrations of 0.2 mM, 1 mM, 5 mM, 10 mM, 20 mM, and 40 mM. Wild-type and ubr-1 animals were exposed to glutamate from the L1 to L4 stages and then assessed after 20 hours on glutamate and IVM plates. The same experimental procedure was used for γ-aminobutyric acid (5 mM, product number: A2129, Sigma Aldrich) and aspartate (5 mM, product number: A6202, Macklin, Shanghai, China).

In parallel, ubr-1 mutants were treated with ceftriaxone sodium (product number: TC6154, Biobomei, Hefei, China) at final concentrations of 0.05 µg/mL, 0.5 µg/mL, 5 µg/mL, 50 µg/mL, and 500 µg/mL. Viability and motor ability were evaluated after L4 worms were transferred to plates containing ceftriaxone and ivermectin (5 ng/mL) for 20 hours.

Molecular biology and RNA interference

All expression and rescue plasmids were generated via the multisite gateway system (50). Three entry clones (Slot1, Slot2, Slot3) corresponding to the promoter, target gene and fluorescent marker genes were recombined into pDEST™ R4-R3 Vector II via the LR reaction, and expression constructs were subsequently obtained. The expression patterns of ubr-1, avr-15, avr-14 and glc-1 were examined via promoter elements of 1.6 kb, 2.7 kb, 3.0 kb and 1.14 kb, respectively. DNA fragments were PCR-amplified from N2 genomic DNA.

Double-stranded RNA (dsRNA) interference technology was used to suppress gene expression in specific tissues. Essential to this process is the expression of dsRNA, which is critical for achieving targeted neuronal or tissue-specific RNAi in C. elegans (51). The plasmid for RNAi was obtained by constructing Slot2 via the BP reaction of a multisite gateway system. The exon-rich region is usually selected as the interfering region (500-700 bp), the sense strand is selected as the target gene for forward interfering sequences (RNAi, sense), and the antisense strand is selected as the target gene for reverse interfering sequences (RNAi, antisense).

Detailed plasmid and primer information is available in Tables S2-S3.

RNA sequencing analysis

For RNA-seq analysis, synchronized L4 worms were washed with bath solution at least 3 times and harvested via centrifugation. Total RNA was isolated via the freeze‒thaw method, and different samples were provided to Novogene for RNA-seq analysis. Transcripts per million (TPM), a standard gene expression normalization method that adjusts the expression of a gene to the number of transcripts per million (52), was used to indicate quantitative gene expression. The TPM value takes into account the length of the gene and the sequencing depth and converts the expression of a gene to the number of transcripts per million by dividing the count value of each gene by its length and normalizing appropriately.

Fluorescence microscopy

L4 stage worms were immobilized with 2.5 mM levamisole (Sigma‒Aldrich, USA) on thin layers of agarose for imaging with a laser scanning confocal microscope (Olympus FV3000, Japan) with a 60× oil objective (numerical aperture = 1.35). The intensities of AChR and GABAAR in strains EN208 kr208 [Punc-29::UNC-29::tagRFP] and KP5931 nuIs283 [Pmyo-3::UNC-49::GFP] were measured. ImageJ software was used to process the images and perform intensity analysis, and the results are reported as the normalized fluorescence intensity per individual and are shown as the mean ± SEM.

In situ Ca2+ imaging

The strain SGA616 [Pmyo-2::GCaMP6s::wCherry] was used to facilitate calcium imaging of the pharyngeal muscles. One-day-old adult hermaphrodites were glued, and the terminal bulb was dissected for imaging with a 60x water objective (Nikon, Japan; numerical aperture = 1.0). Serotonin (5-HT, product number: S31021, Yuanye, Shanghai, China) and ivermectin stock solutions were diluted to the desired concentrations via the following bath solution (in mM): NaCl 150; KCl 5; CaCl2 5;

MgCl2 1; glucose 10; sucrose 5; and HEPES 15, pH 7.3 with NaOH, ∼330 mOsm. The control bath solution, 5-HT alone, or a combination of 5-HT and ivermectin were administered 30 seconds after initiating the recording via a gravity-driven perfusion system (INBIO MPS-3, China). Fluorescence images were acquired with an excitation wavelength LED at 470 nm and a digital sCMOS camera (Hamamatsu ORCA-Flash 4.0V2) under a wide-field microscope (Nikon LV-TV) at a rate of 10 frames per second for 2 minutes. Data were collected from HCImage (Hamamatsu) and analyzed via Image-Pro Plus 6.0 (Media Cybernetics, Inc., Rockville, MD, USA) and ImageJ (National Institutes of Health). The fluorescence intensity of the region of interest (ROI) was defined as F, and the background intensity near the ROI was defined as F0. The true muscle calcium fluorescence signal was obtained by subtracting the background signal from the ROI. ΔF/F0 = (F - F0)/F0 was plotted over time as a fluorescence variation curve.

Statistical analysis

GraphPad Prism 8 was used for statistical analyses and graphing. Dose‒response residual ratio (DRR), IC50, and EC50 values were determined via a sigmoidal dose‒ response model. Unpaired Student’s t test was used to analyze data between two groups, and ordinary one-way or two-way ANOVA was performed for statistical analysis when multiple groups of data were compared. The p values are indicated as follows: ns, not significant, *p < 0.05, **p < 0.01, ***p < 0.001, ****p < 0.0001. The error bars represent the SEMs.

eLife Assessment

This important study will allow for a better understanding of anthelmintic drug resistance in nematodes, a growing concern for public health. The authors provide a detailed analysis of the role of UBR-1 and its underlying mechanism in ivermectin resistance using convincing behavioural and genetic experiments with C. elegans. What is not yet clear is how much of this study applies to parasitic nematodes, in which resistance has naturally emerged in different populations. The study will be of relevance to colleagues working on the evolution of drug resistance and to parasitologists in general.

Reviewer #1 (Public review):

Summary:

The drug Ivermectin is used to effectively treat a variety of worm parasites in the world, however resistance to Ivermectin poses a rising challenge for this treatment strategy. In this study, the authors found that loss of the E3 ubiquitin ligase UBR-1 in the worm C. elegans results in resistance to Ivermectin. In particular, the authors found that ubr-1 mutants are resistant to the effects of Ivermectin on worm viability, body size, pharyngeal pumping, and locomotion. The authors previously showed that loss of UBR-1 disrupts homeostasis of the amino acid and neurotransmitter glutamate resulting in increased levels of glutamate in C. elegans. Here, the authors found that the sensitivity of ubr-1 mutants to Ivermectin can be restored if glutamate levels are reduced using a variety of different methods. Conversely, treating worms with exogenous glutamate to increase glutamate levels also results in resistance to Ivermectin supporting the idea that increased glutamate promotes resistance to Ivermectin. The authors found that the primary known targets of Ivermectin, glutamate-gated chloride channels (GluCls), are downregulated in ubr-1 mutants providing a plausible mechanism for why ubr-1 mutants are resistant to Ivermectin. Although it is clear that loss of GluCls can lead to resistance to Ivermectin, this study suggests that one potential mechanism to decrease GluCl expression is via disruption of glutamate homeostasis that leads to increased glutamate. This study suggests that if parasitic worms become resistant to Ivermectin due to increased glutamate, their sensitivity to Ivermectin could be restored by reducing glutamate levels using drugs such as Ceftriaxone in a combination drug treatment strategy.

Strengths:

(1) The use of multiple independent assays (i.e., viability, body size, pharyngeal pumping, locomotion, and serotonin-stimulated pharyngeal muscle activity) to monitor the effects of Ivermectin

(2) The use of multiple independent approaches (got-1, eat-4, ceftriaxone drug, exogenous glutamate treatment) to alter glutamate levels to support the conclusion that increased glutamate in ubr-1 mutants contributes to Ivermectin resistance.

Weaknesses:

(1) The primary target of Ivermectin is GluCls so it is not surprising that alteration of GluCl expression or function would lead to Ivermectin resistance.

(2) It remains to be seen what percent of Ivermectin-resistant parasites in the wild have disrupted glutamate homeostasis as opposed to mutations that more directly decrease GluCl expression or function.

Reviewer #2 (Public review):

Summary:

The authors provide a very thorough investigation of the role of UBR-1 in anthelmintic resistance using the non-parasitic nematode, C. elegans. Anthelmintic resistance to macrocyclic lactones is a major problem in veterinary medicine and likely just a matter of time until resistance emerges in human parasites too. Therefore, this study providing novel insight into the mechanisms of ivermectin resistance is particularly important and significant.

Strengths:

The authors use very diverse technologies (behavior, genetics, pharmacology, genetically encoded reporters) to dissect the role of UBR-1 in ivermectin resistance. Deploying such a comprehensive suite of tools and approaches provides exceptional insight into the mechanism of how UBR-1 functions in terms of ivermectin resistance.

Weaknesses:

I do not see any major weaknesses in this study. My only concern is whether the observations made by the authors would translate to any of the important parasitic helminths in which resistance has naturally emerged in the field. This is always a concern when leveraging a non-parasitic nematode to shed light on a potential mechanism of resistance of parasitic nematodes, and I understand that it is likely beyond the scope of this paper to test some of their results in parasitic nematodes.

Reviewer #3 (Public review):

Summary:

Li et al propose to better understand the mechanisms of drug resistance in nematode parasites by studying mutants of the model roundworm C. elegans that are resistant to the deworming drug ivermectin. They provide compelling evidence that loss-of-function mutations in the E3 ubiquitin ligase encoded by the UBR-1 gene make worms resistant to the effects of ivermectin (and related compounds) on viability, body size, pharyngeal pumping rate, and locomotion and that these mutant phenotypes are rescued by a UBR-1 transgene. They propose that the mechanism is resistance is indirect, via the effects of UBR-1 on glutamate production. They show mutations (vesicular glutamate transporter eat-4, glutamate synthase got-1) and drugs (glutamate, glutamate uptake enhancer ceftriaxone) affecting glutamate metabolism/transport modulate sensitivity to ivermectin in wild-type and ubr-1 mutants. The data are generally consistent with greater glutamate tone equating to ivermectin resistance. Finally, they show that manipulations that are expected to increase glutamate tone appear to reduce expression of the targets of ivermectin, the glutamate-gated chloride channels, which is known to increase resistance.

There is a need for genetic markers of ivermectin resistance in livestock parasites that can be used to better track resistance and to tailor drug treatment. The discovery of UBR-1 as a resistance gene in C. elegans will provide a candidate marker that can be followed up in parasites. The data suggest Ceftriaxone would be a candidate compound to reverse resistance.

Strengths:

The strength of the study is the thoroughness of the analysis and the quality of the data. There can be little doubt that ubr-1 mutations do indeed confer ivermectin resistance. The use of both rescue constructs and RNAi to validate mutant phenotypes is notable. Further, the variety of manipulations they use to affect glutamate metabolism/transport makes a compelling argument for some kind of role for glutamate in resistance.

Weaknesses:

The proposed mechanism of ubr-1 resistance i.e.: UBR-1 E3 ligase regulates glutamate tone which regulates ivermectin receptor expression, is broadly consistent with the data but somewhat difficult to reconcile with the specific functions of the genes regulating glutamatergic tone. Ceftriaxone and eat-4 mutants reduce extracellular/synaptic glutamate concentrations by sequestering available glutamate in neurons, suggesting that it is extracellular glutamate that is important. But then why does rescuing ubr-1 specifically in the pharyngeal muscle have such a strong effect on ivermectin sensitivity? Is glutamate leaking out of the pharyngeal muscle into the extracellular space/synapse? Is it possible that UBR-1 acts directly on the avr-15 subunit, both of which are expressed in the muscle, perhaps as part of a glutamate sensing/homeostasis mechanism?

The use of single ivermectin dose assays can be misleading. A response change at a single dose shows that the dose-response curve has shifted, but the response is not linear with dose, so the degree of that shift may be difficult to discern and may result from a change in slope but not EC50.

Similarly, in Figure 3C, the reader is meant to understand that eat-4 mutant is epistatic to ubr-1 because the double mutant has a wild-type response to ivermectin. But eat-4 alone is more sensitive, so (eyeballing it and interpolating) the shift in EC50 caused by the ubr-1 mutant in a wild type background appears to be the same as in an eat-4 background, so arguably you are seeing an additive effect, not epistasis. For the above reasons, it would be desirable to have results for rescuing constructs in a wild-type background, in addition to the mutant background.

The added value of the pumping data in Figure 5 (using calcium imaging) over the pump counts (from video) in Figure 1G, Figure 2E, F, K, & Figure 3D, H is not clearly explained. It may have to do with the use of "dissected" pharynxes, the nature/advantage of which is not sufficiently documented in the Methods/Results.

Author response:

The following is the authors’ response to the original reviews.

Public Reviews:

Reviewer #1 (Public review):

Summary:

The drug Ivermectin is used to effectively treat a variety of worm parasites in the world, however resistance to Ivermectin poses a rising challenge for this treatment strategy. In this study, the authors found that loss of the E3 ubiquitin ligase UBR-1 in the worm C. elegans results in resistance to Ivermectin. In particular, the authors found that ubr-1 mutants are resistant to the effects of Ivermectin on worm viability, body size, pharyngeal pumping, and locomotion. The authors previously showed that loss of UBR-1 disrupts homeostasis of the amino acid and neurotransmitter glutamate resulting in increased levels of glutamate in C. elegans. Here, the authors found that the sensitivity of ubr-1 mutants to Ivermectin can be restored if glutamate levels are reduced using a variety of different methods. Conversely, treating worms with exogenous glutamate to increase glutamate levels also results in resistance to Ivermectin supporting the idea that increased glutamate promotes resistance to Ivermectin. The authors found that the primary known targets of Ivermectin, glutamate-gated chloride channels (GluCls), are downregulated in ubr-1 mutants providing a plausible mechanism for why ubr-1 mutants are resistant to Ivermectin. Although it is clear that loss of GluCls can lead to resistance to Ivermectin, this study suggests that one potential mechanism to decrease GluCl expression is via disruption of glutamate homeostasis that leads to increased glutamate. This study suggests that if parasitic worms become resistant to Ivermectin due to increased glutamate, their sensitivity to Ivermectin could be restored by reducing glutamate levels using drugs such as Ceftriaxone in a combination drug treatment strategy.

Strengths:

(1) The use of multiple independent assays (i.e., viability, body size, pharyngeal pumping, locomotion, and serotonin-stimulated pharyngeal muscle activity) to monitor the effects of Ivermectin

(2) The use of multiple independent approaches (got-1, eat-4, ceftriaxone drug, exogenous glutamate treatment) to alter glutamate levels to support the conclusion that increased glutamate in ubr-1 mutants contributes to Ivermectin resistance.

Weaknesses:

(1) The primary target of Ivermectin is GluCls so it is not surprising that alteration of GluCl expression or function would lead to Ivermectin resistance.

(2) It remains to be seen what percent of Ivermectin-resistant parasites in the wild have disrupted glutamate homeostasis as opposed to mutations that more directly decrease GluCl expression or function.

Thank you for your thoughtful and constructive comments. We completely agree with your observation that alterations in GluCl expression or function can lead to Ivermectin resistance. However, we would like to emphasize that our study highlights an additional mechanism: disruptions in glutamate homeostasis can also lead to decreased GluCl expression, thereby contributing to Ivermectin resistance. This mechanism, which has not been fully explored previously, offers new insights into the complexity of drug resistance and could have important implications for understanding the development of Ivermectin resistance in parasitic nematodes.

As you pointed out, the role of disrupted glutamate homeostasis in wild parasitic populations and the proportion of resistant parasites with this mechanism remain unknown. We believe this uncertainty underlines the significance of our findings, as they suggest a novel avenue for studying Ivermectin resistance and for developing potential strategies to counteract it.

We have incorporated this discussion into the revised manuscript to further enrich the context of our findings.

Reviewer #2 (Public review):

Summary:

The authors provide a very thorough investigation of the role of UBR-1 in anthelmintic resistance using the non-parasitic nematode, C. elegans. Anthelmintic resistance to macrocyclic lactones is a major problem in veterinary medicine and likely just a matter of time until resistance emerges in human parasites too. Therefore, this study providing novel insight into the mechanisms of ivermectin resistance is particularly important and significant.

Strengths:

The authors use very diverse technologies (behavior, genetics, pharmacology, genetically encoded reporters) to dissect the role of UBR-1 in ivermectin resistance. Deploying such a comprehensive suite of tools and approaches provides exceptional insight into the mechanism of how UBR-1 functions in terms of ivermectin resistance.

Weaknesses:

I do not see any major weaknesses in this study. My only concern is whether the observations made by the authors would translate to any of the important parasitic helminthes in which resistance has naturally emerged in the field. This is always a concern when leveraging a non-parasitic nematode to shed light on a potential mechanism of resistance of parasitic nematodes, and I understand that it is likely beyond the scope of this paper to test some of their results in parasitic nematodes.

Thank you for your kind words and positive feedback on our work. We greatly appreciate your acknowledgment of the diverse technologies and comprehensive approaches we utilized to uncover the role of UBR-1 in ivermectin resistance.

Your concern about whether our findings in C. elegans translate to parasitic helminthes in which ivermectin resistance has naturally emerged is both valid and critical. This is indeed a key question we expect to figure out in future studies. Collaborating with parasitologists to investigate whether naturally occurring mutations in ubr-1 exist in parasitic and non-parasitic nematodes is a priority for us. We hope that these efforts will lead to meaningful discoveries that have a significant impact on both livestock management and medicine.

Reviewer #3 (Public review):

Summary:

Li et al propose to better understand the mechanisms of drug resistance in nematode parasites by studying mutants of the model roundworm C. elegans that are resistant to the deworming drug ivermectin. They provide compelling evidence that loss-of-function mutations in the E3 ubiquitin ligase encoded by the UBR-1 gene make worms resistant to the effects of ivermectin (and related compounds) on viability, body size, pharyngeal pumping rate, and locomotion and that these mutant phenotypes are rescued by a UBR-1 transgene. They propose that the mechanism is resistance is indirect, via the effects of UBR-1 on glutamate production. They show mutations (vesicular glutamate transporter eat-4, glutamate synthase got-1) and drugs (glutamate, glutamate uptake enhancer ceftriaxone) affecting glutamate metabolism/transport modulate sensitivity to ivermectin in wild-type and ubr-1 mutants. The data are generally consistent with greater glutamate tone equating to ivermectin resistance. Finally, they show that manipulations that are expected to increase glutamate tone appear to reduce expression of the targets of ivermectin, the glutamate-gated chloride channels, which is known to increase resistance.

There is a need for genetic markers of ivermectin resistance in livestock parasites that can be used to better track resistance and to tailor drug treatment. The discovery of UBR-1 as a resistance gene in C. elegans will provide a candidate marker that can be followed up in parasites. The data suggest Ceftriaxone would be a candidate compound to reverse resistance.

Strengths:

The strength of the study is the thoroughness of the analysis and the quality of the data. There can be little doubt that ubr-1 mutations do indeed confer ivermectin resistance. The use of both rescue constructs and RNAi to validate mutant phenotypes is notable. Further, the variety of manipulations they use to affect glutamate metabolism/transport makes a compelling argument for some kind of role for glutamate in resistance.

Weaknesses:

The proposed mechanism of ubr-1 resistance i.e.: UBR-1 E3 ligase regulates glutamate tone which regulates ivermectin receptor expression, is broadly consistent with the data but somewhat difficult to reconcile with the specific functions of the genes regulating glutamatergic tone. Ceftriaxone and eat-4 mutants reduce extracellular/synaptic glutamate concentrations by sequestering available glutamate in neurons, suggesting that it is extracellular glutamate that is important. But then why does rescuing ubr-1 specifically in the pharyngeal muscle have such a strong effect on ivermectin sensitivity? Is glutamate leaking out of the pharyngeal muscle into the extracellular space/synapse? Is it possible that UBR-1 acts directly on the avr-15 subunit, both of which are expressed in the muscle, perhaps as part of a glutamate sensing/homeostasis mechanism?

Thank you for your insightful feedback and thought-provoking questions. These are excellent points that have prompted us to critically reconsider our findings and the proposed mechanism.

Several potential explanations could be considered, although we currently lack direct evidence to support this hypothesis: (1) The pharynx likely plays a dominant role in ivermectin resistance, as previously reported (Dent et al., 1997; Dent et al., 2000), and overexpression of UBR-1 in the pharyngeal muscle may exhibit a strong effect on ivermectin sensitivity. (2) It is also possible that pharyngeal muscle cells have the capacity to release glutamate into the extracellular space, which could contribute to the observed effect. (3) Alternatively, UBR-1 expression in the pharyngeal muscle may regulate other indirect pathways affecting extracellular or synaptic glutamate concentrations.

We also appreciate your suggestion that UBR-1 may act directly on AVR-15 in the pharynx. While this is an interesting possibility, UBR-1 is an E3 ubiquitin ligase, and if AVR-15 were a direct target, we would expect UBR-1 to ubiquitinate AVR-15 and promote its degradation. In this case, loss of UBR-1 should inhibit AVR-15 ubiquitination, reduce its degradation, and lead to increased AVR-15 protein levels in the pharynx. However, our experimental data show a reduction, rather than an increase, in AVR-15::GFP levels in ubr-1 mutants (Figure 4A). This observation suggests that AVR-15 is less likely to be a direct target of UBR-1. To definitively address this hypothesis, a direct assessment of AVR-15 ubiquitination levels in wild-type and ubr-1 mutant backgrounds would be needed. We agree that this is an important avenue for future investigation.

The use of single ivermectin dose assays can be misleading. A response change at a single dose shows that the dose-response curve has shifted, but the response is not linear with dose, so the degree of that shift may be difficult to discern and may result from a change in slope but not EC50. Similarly, in Figure 3C, the reader is meant to understand that eat-4 mutant is epistatic to ubr-1 because the double mutant has a wild-type response to ivermectin. But eat-4 alone is more sensitive, so (eyeballing it and interpolating) the shift in EC50 caused by the ubr-1 mutant in a wild type background appears to be the same as in an eat-4 background, so arguably you are seeing an additive effect, not epistasis. For the above reasons, it would be desirable to have results for rescuing constructs in a wild-type background, in addition to the mutant background.

Thank you for your detailed feedback and observations.

The potential additive effect you noted in Figure 3C appears to be specific to the body length analysis. In our other three ivermectin resistance assays (viability, pumping rate, and locomotion velocity), this additive effect was not observed. A possible explanation for this is that eat-4 and got-1 single mutants inherently exhibit reduced body length compared to wild-type worms (Mörck and Pilon 2006; Greer et al. 2008; Chitturi et al. 2018), which may give the appearance of an additive effect in this particular assay.

Regarding the use of rescuing constructs, we performed these experiments in the ubr-1;got-1 and ubr-1;eat-4 double mutant backgrounds. This was designed to test whether the suppression of ubr-1-mediated ivermectin resistance by got-1 or eat-4 mutations is indeed due to the functional activity of GOT-1 and EAT-4, respectively. The choice of this setup was to ensure that the double mutant phenotype was fully addressed. In contrast, rescuing constructs of GOT-1 and/or EAT-4 in a wild-type background might not sufficiently reveal the relationship between GOT-1, EAT-4, and UBR-1. However, we are open to further testing your suggestion in the future.

To aid in the interpretation and clarify the apparent effects, we have revised Figure 3 annotation to clearly represent the data and the comparisons being made. We hope this adjustment makes the results more straightforward and easier for readers to understand.

The added value of the pumping data in Figure 5 (using calcium imaging) over the pump counts (from video) in Figure 1G, Figure 2E, F, K, & Figure 3D, H is not clearly explained. It may have to do with the use of "dissected" pharynxes, the nature/advantage of which is not sufficiently documented in the Methods/Results.

Thank you for pointing this out. The behavioral pumping data in Figure 1G, Figure 2E, F, K, & Figure 3D and calcium imaging data in Figure 5 were obtained under different experimental conditions. Specifically, the behavioral assays (pumping rate) were conducted on standard culture plates with freely moving worms, whereas the calcium imaging experiments were performed in a liquid environment with immobilized worms. In the calcium imaging setup, the dissection refers to gently puncturing the epidermis behind head of the worm with a glass electrode to relieve internal pressure, which aids in stabilizing the calcium imaging process and ensures better visualization of pharyngeal muscle activity.

We compared the pharyngeal muscle activity of worms that were not subjected to puncturing the epidermis and found no significant difference when activated by 20 mM serotonin. Therefore, we speculate that there is no direct interaction between the bath solution and the pharynx or head neurons. To avoid confusion, we have removed the term "dissected" from the manuscript and added additional experimental details in the Methods section.

Recommendations for the authors:

Reviewer #1 (Recommendations for the authors):

(1) The authors propose that ubr-1 mutants are resistant to ivermectin due to persistent elevation of glutamate that leads to a compensatory reduction in GluCl levels and thus resistance to Ivermectin. This model would be strengthened by experiments more directly connecting glutamate, GluCls and Ivermectin sensitivity. For example, does overexpression of a relevant GluCl such as AVR-15 restore Ivermectin sensitivity to ubr-1 mutants? Does Ceftriaxone treatment affect the Ivermectin resistance of worms lacking the relevant GluCls (i.e., avr-15, avr-14 and glc-1)? - The model suggests that Ceftriaxone treatment would have no effect in the latter case.

Thank you for your valuable suggestion. Based on your recommendation, we have performed two additional experiments to strengthen our model. First, we conducted an overexpression experiment of AVR-15 and found that it significantly, though partially, restored ivermectin sensitivity in ubr-1 mutants (p < 0.01, Supplemental Figure S5D). Second, we tested the effect of Ceftriaxone treatment on the IVM resistance of avr-15; avr-14; glc-1 triple mutants, which encode the most critical glutamate receptors involved in IVM sensitivity. As expected, we found that Ceftriaxone treatment did not alter the IVM resistance in these triple mutants (Supplemental Figure S5E), supporting the idea that these specific GluCls are key to the observed resistance.

These two experiments provide further support for our proposed model. We have integrated the results into the manuscript, updating the Results section and Supplemental Figure S5D, E, as well as the corresponding Figure Legends.

(2) Line 211 - Ceftriaxone is known to upregulate EAAT2 expression in mammals. Do the authors know if the drug also increases EAAT expression in C. elegans?

Thank you for raising this point. To our knowledge, this is the first study to demonstrate the antagonistic effect of ceftriaxone on ivermectin resistance in C. elegans, particularly in the context of ubr-1-mediated resistance. Ceftriaxone enhances glutamate uptake by increasing the expression of excitatory amino acid transporter-2 (EAAT2) in mammals (Rothstein et al., 2005, Lee et al., 2008). C. elegans has six glutamate transporters encoded by glt-1 and glt-3–7 (Mano et al. 2007).

Compared to testing whether ceftriaxone increases the expression of these EAATs in C. elegans, identifying which specific glt gene targeted by ceftriaxone may better reveal its mechanism of action. To investigate this, we performed a genetic analysis. In the ubr-1 mutant, we individually deleted the six glt genes and found that ceftriaxone’s ability to restore ivermectin sensitivity was specifically suppressed in the ubr-1; glt-1 and ubr-1; glt-5 double mutants (Author response image 1A). This suggests that glt-1 and glt-5 may be the targets of ceftriaxone in C. elegans. In contrast, ivermectin sensitivity was unaffected in the individual glt mutants (Author response image 1B), indicating that a single glt deletion may not be sufficient to alter glutamate level or induce GluRs downregulation. Further studies are needed to determine whether ceftriaxone directly increases GLT-1 and GLT-5 expression in C. elegans and to explore the underlying mechanisms.

Glutamate transporter removal inhibits ceftriaxone-mediated restoration of ivermectin sensitivity in ubr-1.

(A) Compared to the ubr-1 mutants, the ubr-1; glt-1 and ubr-1; glt-5 double mutants show enhanced ivermectin resistance under ceftriaxone treatment. (B) The glt mutants do not show resistance to ivermectin. ****p < 0.0001; one-way ANOVA test.

(3) Line 64 - as part of the rationale for the study, the authors state that "...increasing reports of unknown causes of IVM resistance continue to emerge...suggesting that additional unknown mechanisms are awaiting investigation." While this may be true, the ultimate conclusion from this study is that decreasing expression of Ivermectin-targeted GluCls causes Ivermectin resistance, which is a known mechanism. The field already knows that Ivermectin targets GluCls and thus decreasing GluCl expression or function would lead to Ivermectin resistance. The authors may want to edit the sentence mentioned above for clarity.

Thanks for the suggestion. We have revised the sentence for clarity: “…, suggesting that previously unrecognized or additional mechanisms regulating GluCls expression may await further investigation.” This revision better reflects the focus on GluCl regulation and clarifies the potential for additional mechanisms to be explored.

(4) The introduction to the serotonin-stimulated pharyngeal Calcium imaging section is a little confusing. The role of the various GluCls in pharyngeal pumping should be defined/clarified in the introduction to the last section (lines 337-341).

Thanks. We have revised and clarified the introduction as follows: “GluCls downregulation was functionally validated by the diminished IVM-mediated inhibition of serotonin-activated pharyngeal Ca2+ activity observed in ubr-1 mutants. ”

Additionally, the role of the various GluCls in pharyngeal pumping has been clarified:

“Using translational reporters, we found that IVM resistance in ubr-1 mutants is caused by the functional downregulation of IVM-targeted GluCls, including AVR-15, AVR-14, and GLC-1. These receptors are activated by glutamate to facilitate chloride ion influx into pharyngeal muscle cells, resulting in the inhibition of muscle contractions and the suppression of food intake in C. elegans. ”

We hope these revisions address the concerns raised and improve the clarity of this section.

(5) The color code key on the right-hand side of the Raster Plots in Figure 1H should be made larger for clarity.

Revised.

(6) In Figure S3, a legend should be included to define the black and blue box plots.

Thank you for your comment. We have added the following clarification to the figure legend: “Black plots: wild-type, blue plots: ubr-1 mutants.” This should now make the distinction between the two groups clear.

(7) Figure S4, the brackets above the graphs are misleading. It is not clear which comparisons are being made.

Thank you for your feedback. We have clarified the figure by updating the legend to include the statement: “All statistical analyses were performed against the ubr-1 mutant.” This clarification is now also included in Figure 3F-I to ensure consistency and avoid any confusion regarding the comparisons being made.

Reviewer #2 (Recommendations for the authors):

(1) In Figure 1A: the "trails" table needs more clarification to orient the reader.

To improve clarity and better orient the reader, we have updated Figure 1A by explicitly adding the number of trials and including a statistical analysis of the viability of wild-type and ubr-1 mutants under different ML conditions. In Figure 1A legend, we have added “we used shades of red to represent worm viability on each experimental plate (n = 50 animals per plate), with darker shades indicating lower survival rates. The viability test was repeated at least 5 times (5 trials).”. These modifications aim to provide a clearer understanding of the data presentation and its significance.

(2) In Figure S2: it would benefit the reader to include the major human parasitic nematodes in the phylogeny and include a discussion of the conservation.

Thank you for your insightful comment. In Figure S2A, we have included the human parasitic nematodes Onchocerca volvulus, Brugia malayi, and Toxocara canis. Unfortunately, other major human parasitic nematodes, such as Ascaris lumbricoides (roundworm), Ancylostoma duodenale (hookworm), and Trichuris trichiura (whipworm), currently lack reported homologs of the ubr-1 gene.

To provide some context, Onchocerca volvulus is a leading cause of infectious blindness globally, affecting millions of people, while Brugia malayi causes lymphatic filariasis, a significant tropical disease. Toxocara canis is a zoonotic parasite responsible for serious human syndromes such as visceral and ocular larval migration. Ivermectin remains a primary treatment for these parasitic infections.

Interestingly, while we have identified relevant sequences in Onchocerca volvulus, Brugia malayi, and Toxocara canis, potential mutations in ubr-1-like genes in these parasitic nematodes may lead to ivermectin resistance. Sequence comparison analysis could shed light on the risks of such mutations and their relevance to ivermectin treatment failure, warranting further attention. We have added a discussion of this potential risk in the manuscript.

Reviewer #3 (Recommendations for the authors):

Minor corrections/suggestions:

(1) The level of resistance in ubr-1 is similar to dyf genes. Should double-check ubr-1 mutant is not dyf.

Thank you for your insightful suggestion. We are also interested in this point and designed the following experiments. We first directly tested the Dyf phenotype of ubr-1 using standard DIO dye staining (Author response image 2A) and found that ubr-1 clearly show a "dye filling defective" phenotype (Author response image 2B). This raises an interesting question: Could the IVM resistance observed in ubr-1 be due to its Dyf defect? To address this, we further performed experiment by using Ceftriaxone to test ubr-1’s Dyf phenotype. Ceftriaxone can fully rescue the sensitivity of ubr-1 to IVM (Figure 2). If IVM resistance observed in ubr-1 is due to its Dyf defect, we should observe same rescued Dyf defect. After treating ubr-1 mutants with Ceftriaxone (50 μg/mL) until L4 stage, we again performed DIO dye staining and found that while Ceftriaxone fully rescued IVM resistance in ubr-1, it did not rescue the Dyf defect (Author response image 2C). These results suggest that while ubr-1 has a Dyf defect, it is unlikely the primary cause of the IVM resistance in ubr-1 mutant.

ubr-1 mutant is not dyf. (A) Depiction of the DIO dye-staining assays.

Diagram is adapted from (Power et al. 2020). (B) ubr-1 mutant exhibits obvious Dyf phenotype. (C) Cef treatment (50 μg/mL) does not alter the ubr-1 Dyf defect phenotype. Scale bar, 20 µm.

(2) 367 "in IVM" superscript.

(3) 429 ubr-1 italics.

Thanks, revised.

(4) Methods: Need more info on dissection: if there is direct interaction of bath with pharynx, as suggested by bath solution, then 5HT concentrations are too high. Direct exposure to 20mM 5HT will kill a pharynx. 20uM 5HT?

Thank you for your comment. We have reviewed our experimental records and confirmed that the concentration mentioned in the manuscript is correct. In our experiment, the dissection refers to gently puncturing the epidermis behind head of the worm with a glass electrode to relieve internal pressure, which helps stabilize the calcium imaging process. In fact, there is no direct interaction between the bath solution and the pharynx or head neurons. We have revised the Methods section to clarify this point.

(5) Figure 2: Meaning of "Trials" arrow on grid y-axis is not immediately obvious to me. Would prefer you just label/number individual trials.

Sure, we have labeled the trails accordingly in revised Figure 1, 2, and Figure S1.

(6) Figure 3: Legend should include [IVM]. Meaning of +EAT-4, +GOT-1 should be described in the legend.

Thank you for your suggestion. We have updated the figure legend to include the IVM concentration (5 ng/mL). Additionally, we have clarified the meaning of +EAT-4 and +GOT-1 in the legend with the description: “…whereas the re-expression of GOT-1 (+GOT-1) and EAT-4 (+EAT-4) partially reinstated IVM resistance in the respective double mutants.” This ensures the figure is more informative and accessible to the reader.

(7) 784 signalling pathway should just be pathway.

Thanks, revised.

(8) Line 811 " Both types of motor neurons are innervated by serotonin (5 -HT)." Innervated by serotonergic "neurons"? However, even that is misleading because serotonin is not necessarily synaptic.

Thank you for your comment. We have revised the sentence to: “Both types of motor neurons could be activated by serotonin (5-HT).” This clarification better reflects the role of serotonin in modulating motor neuron activity.

(9) Line 814 puffing or perfusion. Perfusion seems more accurate. Make the figure consistent.

Thanks, revised.

(10) Figure S1 requires an x axis label with better explanation.

Thank you for your feedback. We have revised Figure S1 and added "x-axis" to clarify that it represents the trail number. Additionally, we have updated the figure legend to include the experimental conditions: “The shades of red represent worm viability, with darker shades indicating lower survival rates, based on 100 animals per plate and at least 5 trials.”

(11) Figure S2 C-F needs ivermectin concentration.

(12) Line 865 plants -> plates?

Thanks, revised.

(13) Figure S4. 875 "Rescue of IVM sensitivity of the ubr-1 mutant by the UBR-1 genomic fragment." Wrong title? Describes GFP expression and RNAi experiments.

Thank you for pointing out the mistake in the title. We have revised the title to: “Knockdown of UBR-1 induces IVM resistance phenotypes.” Additionally, we have updated the figure description to include details about GFP expression and RNAi experiments. The GFP expression is now described as: “Expression of functional UBR-1::GFP, driven by its endogenous promoter, was observed predominantly in the pharynx, head neurons, and body wall muscles with weaker expression detected in vulval muscles and the intestine.” The RNAi experiments are described as: “Double-stranded RNA (dsRNA) interference was employed to suppress gene expression in specific tissues (Methods).”

Version 2

Abstract

Abstract

Resistance to anthelmintics, particularly the macrocyclic lactone ivermectin (IVM), presents a substantial global challenge for parasite control. We found that the functional loss of an evolutionarily conserved E3 ubiquitin ligase, UBR-1, leads to IVM resistance in Caenorhabditis elegans. Multiple IVM-inhibiting activities, including viability, body size, pharyngeal pumping, and locomotion, were significantly ameliorated in various ubr-1 mutants. Interestingly, exogenous application of glutamate induces IVM resistance in wild-type animals. The sensitivity of all IVM-affected phenotypes of ubr-1 is restored by eliminating proteins associated with glutamate metabolism or signaling: GOT-1, a transaminase that converts aspartate to glutamate, and EAT-4, a vesicular glutamate transporter. We demonstrated that IVM-targeted GluCls (glutamate-gated chloride channels) are downregulated and that the IVM-mediated inhibition of serotonin-activated pharynx Ca2+ activity is diminished in ubr-1. Additionally, enhancing glutamate uptake in ubr-1 mutants through ceftriaxone completely restored their IVM sensitivity. Therefore, UBR-1 deficiency-mediated aberrant glutamate signaling leads to ivermectin resistance in C. elegans.

Introduction

Anthelmintic resistance (AR) to broad-spectrum antiparasitic macrocyclic lactones (MLs), which include ivermectin (IVM), avermectin (AVM), and doramectin (DOM), has emerged as a critical global concern in veterinary parasites (1–3). The rise of anthelmintic-resistant nematodes poses a pressing issue, impacting not only livestock production but also animal welfare and human health. Overcoming this challenge requires a comprehensive understanding of the molecular mechanisms driving nematode resistance, which, in turn, will facilitate the development of innovative and effective parasite prevention strategies.

The nematode Caenorhabditis elegans (C. elegans), owing to its sensitivity to a majority of anthelmintic drugs and conservation of functional genes, serves as a crucial model organism for identifying genes related to IVM resistance (4–7). Mutations in three C. elegans glutamate-gated chloride channel (GluCl) genes—avr-15, avr-14, and glc-1—have collectively demonstrated strong resistance to ivermectin (8–11). Exogenous expression and crystallization X-ray structural analysis support the premise that these GluCls act as the primary targets of IVM (12, 13). Further identification of IVM resistance-related genes, including P-glycoprotein transporters (pgp-1, pgp-3, pgp-6, pgp-9, pgp-13, and mrp-6), cytochrome oxidases (cyp-14, cyp-34/35), and glutathione S-transferases (gst-4, gst-10) (14–16), has revealed their role in enhancing ivermectin excretion or metabolism. These detoxification genes are thought to be regulated by the ubiquitous transcription factor NHR-8, which encodes a nuclear hormone receptor (17). While these studies have revealed a polygenic mechanism underlying IVM resistance, increasing reports of unknown causes of IVM resistance continue to emerge (18–20), suggesting that previously unrecognized or additional mechanisms regulating GluCls expression may await for further investigation.

UBR1, an E3 ubiquitin ligase and a pivotal component of the ubiquitin-proteasome system, facilitates protein degradation (21–25). UBR1 is implicated in substrate-specific recognition and metabolism, primarily through the N-end rule (26). In humans, loss-of-function mutations in UBR1 result in Johanson-Blizzard syndrome (JBS), a rare autosomal recessive disorder characterized by a spectrum of developmental and neurological symptoms (27, 28). In the nematode Caenorhabditis elegans, a single UBR-1 homolog retains all major conserved functional domains (29). LIN-28 is the sole identified substrate of C. elegans UBR-1, acting as an RNA-binding pluripotency factor crucial for seam cell patterning (30). However, the absence of LIN-28 neither replicates nor ameliorates the movement defects observed in ubr-1 mutants-stiff body bending during backward movement (29).

Our recent research has consistently shown that UBR-1 regulates neural outputs through glutamate homeostasis mechanisms. Loss of UBR-1 leads to elevated glutamate levels in animals, disrupting coordinated motor patterns, likely due to a compensatory decrease in the expression of excitatory glutamate receptors (29). Furthermore, ubr-1 mutations disturb inherent asymmetric neural activity by activating inhibitory GluCls (GLC-3 and GLC-2/4), which in turn impairs rhythmic defecation motor programs (31). These observations suggest that aberrant glutamate metabolism resulting from UBR-1 deficiency has a widespread effect on glutamatergic signaling pathways. This leads to the question of whether ubr-1 mutants cause comprehensive changes in IVM-targeted GluCls, thereby contributing to IVM resistance.

To investigate this possibility, we assessed the IVM sensitivity of wild-type and ubr-1 mutant animals. We found that, in contrast to wild-type N2 animals, loss-of-function ubr-1 mutants present a range of IVM-resistant phenotypes. N2 animals exhibit high fatality rates, abnormal development and motor defects, such as pharyngeal pumping or locomotion disabilities, upon exposure to ivermectin. In contrast, ubr-1 mutants were almost insensitive to IVM. The elimination of the glutamate synthetase GOT-1 and the glutamate transporter EAT-4 completely reversed the IVM sensitivity of ubr-1 mutants, highlighting the critical role of the glutamatergic signaling pathway. Using translational reporters, we found that IVM resistance in ubr-1 mutants is caused by the functional downregulation of IVM-targeted GluCls, including AVR-15, AVR-14, and GLC-1. These receptors are activated by glutamate to facilitate chloride ion influx into pharyngeal muscle cells, thereby inhibiting muscle contractions and suppressing food intake in C. elegans. This downregulation of GluCls is presumably triggered by the aberrantly elevated glutamate-induced compensatory decrease, as the application of exogenous glutamate in wild-type animals partially replicated the IVM resistance observed in ubr-1 mutants. GluCls downregulation is functionally validated, as evidenced by the diminished IVM-mediated inhibition of serotonin-activated pharyngeal Ca2+ activity in ubr-1 mutants. Moreover, pharmacologically reducing glutamate levels via the use of ceftriaxone (32, 33) successfully restored IVM sensitivity in ubr-1 mutants. Thus, our study presents multiple lines of evidence for a novel IVM resistance mechanism linked to the deficiency of the ubiquitin ligase UBR-1 via a glutamate metabolism pathway.

Results

Dose- and time-dependence of IVM resistance in ubr-1 mutants

To further elucidate the mechanisms of anthelmintic resistance (AR), we employed a simplified assay derived from prior research (Figure 1A). We assessed the survival of nematodes on media infused with antiparasitic macrocyclic lactones (MLs). In essence, synchronized L4-stage worms were placed on NGM plates with OP50 bacteria and treated with various MLs, including ivermectin (IVM, 5 ng/mL), avermectin (AVM, 5 ng/mL), or doramectin (DOM, 5 ng/mL). Twenty hours postexposure, we quantified the surviving worms and calculated the percentage of viability relative to the initial population.

ubr-1 exhibits IVM-resistant phenotypes.

(A) Left: Schematic representation of the IVM resistance test in C. elegans. Well-fed L4 stage animals were transferred to plates containing OP50 bacteria with or without macrocyclic lactones (MLs) for 20 hours. Right: Representative grid plots illustrating the viability of wild-type and ubr-1 animals in response to ivermectin (IVM, 5 ng/mL), avermectin (AVM, 5 ng/mL), and doramectin (DOM, 5 ng/mL). We used shades of red to represent worm viability on each experimental plate (n = 50 animals per plate), with darker shades indicating lower survival rates. The viability test was repeated at least 5 times (5 trials). (B) Quantification analysis of the viability of wild-type and ubr-1(hp684) mutants exposed to different MLs. ubr-1(hp684) mutants exhibit resistance to various MLs. (C) Dose-response curve depicting the viability of wild-type and ubr-1(hp684) mutants in the presence of varying concentrations of IVM. The IC50 was 5.7 ng/mL for the wild type and 11.6 ng/mL for the ubr-1 mutants. (D) Time-dependent effect of IVM exposure on animal viability. The IC50 values were 20.2 hours for the wild type and 42.1 hours for the ubr-1 mutants. (E) Representative images of worm size in the wild type and ubr-1(hp684) mutants with or without IVM treatment. Scale bar, 50 µm. (F) Quantitative analysis of body length in different genotypes with or without IVM. (G) Quantification of the average pharynx pump number in animals with or without IVM. (H) Raster plots illustrating the locomotion velocity of individual animals in the absence (0 ng/mL) and presence (5 ng/mL) of IVM. n = 10 animals in each group. (I) Quantification of the average velocity in different genotypes with or without IVM treatment. ns, not significant, *p < 0.05, **p < 0.01, ***p < 0.001 by Student’s t test. The error bars represent the SEM.

Our observations revealed a significant decline in the viability of wild-type nematodes following ML exposure (IVM 52.0 ± 3.2%, AVM 51.6 ± 2.9%, DOM 52.4 ± 3.4%, in contrast to the vehicle control 100 ± 0.0%), confirming the susceptibility of C. elegans to MLs (Figure 1B). In contrast, ubr-1(hp684; lf) mutants presented remarkable resistance to ML treatments (IVM, ubr-1 96.0 ± 1.7% vs vehicle control 100 ± 0.0%) when juxtaposed with their wild-type counterparts (IVM, WT 52.0 ± 3.2%) (Figure 1B). This trend of increased survival was consistently observed with ubr-1(hp684) mutants treated with AVM (WT 51.6 ± 2.9% vs ubr-1 82.8 ± 2.9%) and DOM (WT 52.4 ± 3.4% vs ubr-1 83.6 ± 3.1%). These findings reveal the pronounced ML resistance characteristics of ubr-1 loss-of-function mutants.

We further investigated the dose- and time-dependent resistance of ubr-1(hp684) to select IVM as a proxy for MLs because of its efficiency and common application. We tested a spectrum of IVM concentrations ranging from 1 ng/mL to 11 ng/mL (Figure S1A). The data revealed that ubr-1(hp684) mutants exhibited resistance at all tested doses (Figure 1C), resulting in a doubling of the half-maximal inhibitory concentration (IC50) — from 5.7 ng/mL in wild-type N2 worms to 11.6 ng/mL in ubr-1 mutants. Time-dependent IVM resistance was also compared between the wild type and ubr-1 mutants (Figure S1B). The duration of half-maximal inhibition in the wild type was approximately 20.2 hours, which increased to 42.1 hours in the ubr-1(hp684) mutants (Figure 1D). To highlight the disparity in IVM resistance, we utilized an assay with 5 ng/mL IVM for a 20-hour period (indicated by the vertical dashed lines in Figure 1C and D) for subsequent experiments, unless otherwise specified.

Additional IVM resistance phenotypes of ubr-1

In addition to viability, ubr-1 mutants maintained normal growth despite IVM exposure. After 20 hours of IVM, the body size of the wild-type animals significantly decreased (34), resulting in a marked decrease in the middle line body length (control, 1.07 ± 0.02 mm; IVM, 0.62 ± 0.03 mm) (Figure 1E). In contrast, the size of the ubr-1 mutants (control, 1.08 ± 0.02 mm; IVM, 1.00 ± 0.02 mm) was largely preserved, indicating that the ubr-1 mutants developed into typical young adults (Figure 1F). Notably, while IVM exposure nearly halted pharyngeal pumping in wild-type animals (control 247.4 ± 4.1 min-1, IVM 34.6 ± 2.8 min-1), ubr-1 mutants sustained considerable pumping activity (control 254.1 ± 5.0 min-1, IVM 152.9 ± 3.0 min-1) (Figure 1G). Motor functions were similarly resistant to IVM in ubr-1 mutants (Movie S1). The velocity of free movement under IVM treatment was significantly greater in ubr-1 mutants (135.9 ± 7.2 μm/s) than in wild-type animals (23.8 ± 2.1 μm/s) (Figure 1H, I, Figure S1C). No significant differences in viability, body length, pharyngeal pumping, or locomotion speed were detected between ubr-1 and wild-type animals in the absence of IVM treatment.

These results collectively indicate that ubr-1 loss-of-function mutants present a comprehensive spectrum of phenotypes associated with IVM resistance.

Different ubr-1 mutants exhibit consistent IVM resistance

The ubr-1 gene is functionally conserved from yeast to humans. In nematodes or filarial worms, ubr-1 also exhibited high sequence similarity (Figure S2A). The hp684 allele contains a premature stop codon, resulting in the truncation of the final 194 amino acids of UBR-1 (Q1864X) (Figure S2B). To validate functional congruence, we expanded our study to include additional ubr-1 mutant alleles. These alleles, which are genetically predicted to lack one or more critical conserved domains, including hp865, which replaces the entire RING finger domain with an SL2-NLS::GFP, and hp821 hp833 (E34X, E1315X), each introducing stop codons that suggest the absence of all functional domains. All these alleles demonstrated IVM resistance comparable to that of the hp684 allele. Specifically, the viability of hp865 (81.6 ± 3.8%) and hp821 hp833 (76.4 ± 2.6%) increased following IVM treatment (Figure S2C). Similarly, the body lengths of hp865 (0.93 ± 0.02 mm) and hp821 hp833 (0.90 ± 0.02 mm) were not diminished by IVM exposure (Figure S2D). Pharyngeal pumping and spontaneous movement velocities were also significantly greater in hp865 (154.2 ± 3.4 min-1, 100.2 ± 7.4 μm/s) and hp821 hp833 (134.9 ± 3.7 min-1, 86.1 ± 6.9 μm/s) than in the wild type (Figure S2E, F). These findings demonstrate that various ubr-1 mutants exhibit a consistent IVM-resistant phenotype. While we have identified relevant sequences in parasitic nematodes including Onchocerca volvulus, Brugia malayi, and Toxocara canis, potential mutations in ubr-1-like genes in these parasitic nematodes may lead to ivermectin resistance.

Glutamate induces IVM resistance in wild-type worms

Systemic elevation of glutamate levels in ubr-1 mutants has been documented (29, 35, 36), prompting an investigation into whether this increase is the key determinant of the observed resistance to IVM.

To test this hypothesis, we pretreated wild-type N2 animals with glutamate to mimic the excessive glutamate environment in ubr-1 mutants (Figure 2A). During these trials, wild-type larvae were exposed to L-glutamate (20 mM) on NGM plates from the L1 to L4 stages before their transfer to IVM-containing plates. Strikingly, these glutamate-pretreated animals presented a significant increase in IVM resistance (Figure 2B).

Glutamate mimics IVM resistance in wild-type N2 animals, and ceftriaxone fully reverses IVM resistance in ubr-1 mutants.

(A) Modified schematic representation of the pretreatment approach in the IVM resistance test in C. elegans. Synchronized L1 animals were cultured on plates containing OP50 bacteria and pretreated with glutamate (Glu, 5 mM), γ-aminobutyric acid (GABA, 5 mM), or aspartate (Asp, 5 mM) until they reached the L4 stage. Pretreated animals were then transferred to plates containing additional IVM (5 ng/mL) for 20 hours. (B) Representative grid plots illustrating the viability of wild-type and ubr-1 animals pretreated with different glutamate metabolites. (C-F) Quantitative analysis of viability, body length, pharyngeal pump rate, and locomotion velocity in wild-type and ubr-1 mutants following exposure to various glutamate metabolites. Wild-type worms treated with glutamate displayed resistance to IVM. However, glutamate did not affect the IVM resistance of the ubr-1 mutant. ns, not significant, ****p < 0.0001 by one-way ANOVA in C-F. (G) Diagram illustrating the ceftriaxone pretreatment procedure in the IVM resistance test. Synchronized L1 animals were cultured on OP50-fed plates with or without ceftriaxone (50 μg/mL) until the L4 stage. The animals were subsequently transferred to plates seeded with additional IVM (5 ng/mL) for 20 hours. (H) Upper: Representative grid plots depicting the viability of ubr-1 animals at different concentrations of ceftriaxone. Bottom: Dose-response curve illustrating the amount of ceftriaxone required for viability in ubr-1(hp684) mutants (IC50 = 5.1 μg/mL). (I-L) Quantification of viability, body length, pharyngeal pumping, and locomotion velocity in the wild type and ubr-1 mutants in the absence and presence of ceftriaxone. Ceftriaxone fully restored the sensitivity of ubr-1 mutants to IVM, comparable to the levels observed in wild-type N2 animals. ns, not significant, ***p < 0.001 by Student’s t test in I-L. Error bars represent SEM.

Specifically, the viability of these glutamate-exposed animals increased from 52 ± 3.2% to 77.2 ± 1.5% compared with that of their nonexposed counterparts (Figure 2C). A corresponding increase in body size was noted, from 0.62 ± 0.03 mm to 0.76 ± 0.01 mm (Figure 2D). Motor activity similarly improved, with velocities increasing from 23.79 ± 2.1 μm/s to 62.4 ± 7.6 μm/s (Figure 2E-F). These findings clearly demonstrate that the external application of glutamate can induce IVM resistance in wild-type N2 animals. Interestingly, in ubr-1 mutants, pretreatment with glutamate did not further increase IVM resistance, implying potential saturation of the glutamate effect in ubr-1 mutants (Figure 2B-F).

In addition to glutamate, other metabolically related molecules, such as aspartate (Asp), a precursor in glutamate synthesis, and γ-aminobutyric acid (GABA), a product of glutamate metabolism, were also found to be elevated in the ubr-1 mutant (29, 31). However, the cocultivation of wild-type N2 worms with aspartate or GABA did not confer IVM resistance (Figure 2A-F), suggesting the glutamate-specific regulation of IVM resistance. This finding highlights the specificity and sufficiency of glutamate in conferring IVM resistance in a wild-type context.

Enhancing glutamate uptake restores IVM sensitivity in ubr-1

Given the likelihood that elevated glutamate levels underpin IVM resistance in ubr-1, mitigating this glutamate surplus could reverse resistance and reinstate IVM sensitivity. To evaluate this premise, we employed ceftriaxone (Cef) (Figure 2G), which is known for enhancing glutamate uptake by upregulating excitatory amino acid transporter-2 (EAAT2) (32, 33), with the aim of reducing glutamate levels.

The results were striking: the IVM resistance typically observed in ubr-1 mutants was completely counteracted by Cef treatment. While Cef (50 μg/mL) had a negligible effect on wild-type N2 animals, it fully restored IVM sensitivity in ubr-1 mutants (Figure 2H). Moreover, this restoration followed a dose-dependent pattern, with a half-effect concentration of 5.14 μg/mL (Figure 2H). Cef treatment resulted in a decrease in the viability of ubr-1 mutants from 96.0 ± 1.7% to 62.8 ± 2.3% in the presence of IVM (Figure 2I). Additional IVM resistance phenotypes in ubr-1 mutants, such as body size, pharyngeal pumping frequency, and locomotion velocity, were fully ameliorated by Cef (Figure 2J-L).

In line with its classification as a beta-lactam antibiotic, Cef effectively reduced the growth of the worm food source E. coli OP50, resulting in a reduced bacterial lawn on NGM plates (Figure S3A, B). However, simply diminishing the OP50 bacterial quantity on NGM plates without Cef did not restore IVM sensitivity in ubr-1 mutants (Figure S3C-F), suggesting that food scarcity is not a critical factor. Similarly, increasing the number of OP50 bacteria present on NGM plates containing Cef did not affect its ability to restore IVM sensitivity in ubr-1 mutants (Figure S3C-F). These findings indicate that the antimicrobial action of ceftriaxone is not related to its ability to restore IVM sensitivity in ubr-1 mutants.

Hence, these results indicate that pharmacologically reducing glutamate levels in ubr-1 reestablishes IVM sensitivity, leading to the hypothesis that aberrant glutamate metabolism underlies IVM resistance in ubr-1.

Restoration of ubr-1 IVM sensitivity through glutamatergic signaling pathway regulation

Our previous research revealed the involvement of UBR-1 in regulating glutamate metabolism, which is essential for coordinated reversal bending and rhythmic defecation motor programs (29). The absence of UBR-1 also leads to perturbations in the balance between the GABAergic and glutamatergic signaling pathways (31). In C. elegans, IVM sensitivity is marked by the sustained activation of glutamate-gated chloride channels (GluCls) by IVM (11). These glutamate-related findings led us to explore whether the resistance of ubr-1 mutants to IVM could be attributed to alterations in the glutamate pathway.

At glutamatergic synapses, a series of enzymatic reactions link glutamate to neurotransmitter recycling. In C. elegans, glutamate-oxaloacetate transaminases, or GOT enzymes, facilitate amino group conversion between aspartate and α-ketoglutarate (α-KG) to oxaloacetate (OAA) and glutamate (Figure 3A) (37, 38). We began our exploration by assessing IVM sensitivity following the disruption of GOT-1. While single got-1 loss-of-function mutants displayed IVM sensitivity similar to that of the wild type, ubr-1; got-1 double mutants presented a significant increase in IVM sensitivity, reverting to wild-type levels of viability, body length, and motor activity, in stark contrast to ubr-1 single mutants (Figure 3B-E). These findings indicate that impeding glutamate synthesis can effectively restore IVM sensitivity in ubr-1 mutants.

Deletion of got-1 and eat-4 suppresses ubr-1’s IVM resistance.

(A) Schematic of the pathway for glutamate synthesis through transaminase GOT-1 and loading via vesicular transport VGLUT/EAT-4. (B-E) Quantification of viability, body length, pharyngeal pump rate, and locomotion velocity in different genotypes. The removal of got-1 and eat-4 significantly restored sensitivity to IVM (5 ng/mL) in ubr-1 mutants, whereas the re-expression of GOT-1 (+GOT-1) and EAT-4 (+EAT-4) partially reinstated IVM resistance in the respective double mutants. (F-I) IVM resistance phenotypes were rescued by expression of UBR-1 (+UBR-1) driven by its own promoter (Pubr-1) or the pharynx muscle-specific promoter (Pmyo-2). For the viability test in F, n = 50 animals per plate and repeated at least 5 times (5 trials), n ≥ 10 animals in G-I, *p < 0.05, **p < 0.01, ***p < 0.001 by one-way ANOVA. All statistical analyses were performed against ubr-1 mutant. Error bars, SEM.

Glutamate, a crucial excitatory neurotransmitter, is packed into synaptic vesicles by the vesicular glutamate transporter (VGLUT) and released into the synaptic cleft to activate postsynaptic receptors. We next explored the impact of inhibiting glutamate transport in ubr-1 mutants by eliminating EAT-4, the sole VGLUT in C. elegans (39). Notably, the absence of EAT-4 in ubr-1 mutants dramatically reversed their IVM resistance, as evidenced by increased viability (ubr-1; eat-4 66.8 ± 3.5%), reduced body length (ubr-1; eat-4 0.56 ± 0.02 mm), increased pharyngeal pumping (ubr-1; eat-4 46.7 ± 3.6 min-1), and improved locomotion speed (ubr-1; eat-4 83.5 ± 7.4 μm/s) in ubr-1; eat-4 double mutants (Figure 3B-E).

In summary, these results underscore the importance of glutamate synthesis and synaptic transport in restoring IVM sensitivity in ubr-1 mutants, suggesting a role for synaptic glutamate signaling in the resistance of ubr-1 to IVM. The IVM resistance phenotypes of wild-type N2 animals were not rescued by ceftriaxone (Figure 2I-L), suggesting that ceftriaxone appears to be effective only against excess glutamate. This discovery is particularly intriguing, as it reveals a previously unexplored role of glutamate metabolism in the regulation of IVM resistance.

UBR-1 appears to regulate IVM resistance from the pharynx

UBR-1 is ubiquitously expressed across a variety of tissues, including the pharynx, body wall muscles, and neurons (Figure S4A) (29, 35, 36). To pinpoint the critical tissue for the role of UBR-1 in IVM resistance, we conducted tissue-specific rescue assays via a functional plasmid. Our findings revealed that restoring UBR-1 expression under its endogenous promoter attenuated the IVM resistance observed in ubr-1 mutants. Notably, when UBR-1 expression was driven by an exogenous pharyngeal muscle-specific promoter, it fully corrected the IVM resistance phenotypes of the ubr-1 mutants (Figure 3F-I). These results imply that although UBR-1 is important in multiple tissues, its role in the pharynx is particularly critical in modulating IVM sensitivity.

To verify the tissue-specific function of UBR-1, we utilized RNA interference (RNAi) to selectively decrease ubr-1 expression in the pharynx. This precise RNAi application to the pharynx indeed induced IVM resistance in wild-type N2 worms (Figure S4B-E). RNAi-mediated knockdown of ubr-1 in nonpharyngeal tissues, such as neurons and body wall muscles, also elicits an IVM resistance profile. Our research revealed that animals in which ubr-1 was knocked down specifically in the pharynx presented more pronounced IVM resistance phenotypes than did those in which ubr-1 was knocked down in other tissues. This phenotype approached the level observed when ubr-1 was knocked down via its native promoter, underscoring the unique contribution of the pharynx to ubr-1-mediated IVM resistance.

Taken together, these results confirm that UBR-1 predominantly regulates IVM resistance through its action in the pharynx.

Downregulation of IVM-targeted GluCls in ubr-1

What mechanisms underlie the induction of IVM resistance by abnormal glutamate metabolism? Previously, we established that ubr-1 mutants exhibit impaired GLR-1 glutamate receptor function in AVA/AVE interneurons (29). Three GluCls (avr-15, avr-14, glc-1) are recognized as pivotal for IVM resistance in C. elegans (11). We sought to determine whether compromised IVM-targeted GluCls contribute to IVM resistance in ubr-1 mutants.

To test this hypothesis, we created three transgenic strains with translational reporters for each GluCl under the control of their native promoters (Pavr-15::AVR-15::GFP, Pavr-14::AVR-14::GFP, Pglc-1::GLC-1::GFP). Notably, AVR-15::GFP expression was pronounced in the pharynx, whereas GLC-1::GFP displayed a distinct head expression pattern, particularly concentrated in the head body wall muscles (Figure 4A, Figure S5A). AVR-14::GFP expression was localized primarily to various head and body neurons (Figure S5A). Intriguingly, compared with that in wild-type animals, the fluorescence intensity of these GluCls was markedly lower in ubr-1 mutants (Figure S5A, B). The modulation of GluCl expression by ubr-1 seems to be posttranscriptional, given that the RNA levels of the GluCl-encoding genes (avr-15, avr-14, glc-1) did not significantly differ (Figure S5C).

Downregulation of AVR-15 in ubr-1 mutants.

(A) Representative fluorescence intensity of AVR-15 in different genotypes. (B) Quantitative analysis of normalized fluorescence intensity revealed a reduction in AVR-15::GFP expression in ubr-1 mutants. The diminished AVR-15 levels were rescued by overexpression of UBR-1 and genetically suppressed in got-1 and eat-4 mutants. Pharmacological intervention with ceftriaxone (50 μg/mL) successfully reinstated the pharyngeal expression of AVR-15::GFP. Conversely, glutamate administration reduced AVR-15 expression in the wild-type N2 strain. ROI, region of interest; n ≥ 10 animals in each group. **p < 0.01, ****p < 0.0001 by one-way ANOVA. Error bars, SEM.

Crucially, the reduction in pharyngeal AVR-15::GFP fluorescence observed in ubr-1 mutants was reversed upon reintroduction of UBR-1 (Figure 4A). Additionally, eliminating got-1 and eat-4 in these mutants mitigated the decrease in AVR-15::GFP expression within the pharynx (Figure 4A, B). Overexpression of AVR-15 also partially restored the sensitivity of ubr-1 mutant to IVM (Figure S5D). These results suggest a direct role of glutamate signaling in regulating GluCl expression levels. Notably, ceftriaxone pretreatment effectively reversed the diminished pharyngeal AVR-15::GFP levels in ubr-1 mutants (Figure 4A, B). In addition to glutamate pretreatment eliciting IVM resistance in wild-type animals, the application of glutamate markedly diminished the pharyngeal AVR-15::GFP fluorescence in these same wild-type worms (Figure 4A, B). Interestingly, the ability of ceftriaxone pretreatment to rescue viability in ubr-1 mutants was completely lost in the avr-14; avr-15; glc-1 triple mutants (Figure S5E), confirming that ceftriaxone restores IVM sensitivity by reducing glutamate levels, which in turn upregulates GluCl receptors.

To further investigate the potential regulatory influence of ubr-1 on a broader spectrum of synaptic receptors, we selected two distinct categories of receptors that are not activated by glutamate. This reduction was not observed for the excitatory acetylcholine receptor (UNC-29/AChR) or inhibitory GABA receptor (UNC-49/GABAAR) (Figure S6), suggesting that ubr-1 may selectively affect glutamate-gated receptors (29).

Collectively, these results demonstrate that aberrant glutamate metabolism in the ubr-1 mutant leads to the downregulation of IVM-targeted GluCls, which likely contributes to diminished IVM responsiveness, culminating in the IVM resistance observed in ubr-1 mutants.

IVM completely suppresses serotonin-evoked pharynx activity in wild-type animals but not in ubr-1

IVM acts by engaging target GluCls, leading to the irreversible inhibition of pharyngeal pumping, disrupting conduction between nerve and muscle cells and eventually killing parasites (8, 11). The intricate dynamics of pharyngeal pumping in C. elegans rely on the coordinated interaction between the pharyngeal muscles and the excitatory pacemaker neuron MC, which is activated by serotonin (40).

To further examine the functional impairments of IVM-targeted GluCls, we conducted a direct examination of the impact of IVM on pharyngeal muscle cell activity. We generated a transgenic line expressing a Ca2+ sensor, GCaMP6, in the pharynx, enabling real-time monitoring of pharyngeal muscle activity (Figure 5A, Methods). Using 5-HT to stimulate excitatory MC neurons, we induced intrinsic Ca2+ activity within the pharyngeal muscles. In these assays, 5-HT triggered strong Ca2+ transients across the pharyngeal muscle cells, from the terminal bulb through the isthmus to the anterior bulb (Figure 5A). To align with our quantified pumping behavior analysis, we focused our observations on the Ca2+ intensity within the terminal bulb. We noted a dose-dependent response of pharyngeal muscle cells to 5-HT, with an estimated EC50 of approximately 20.1 mM (Figure 5B, C). This assay provides a practical method to evaluate the functional impairment of IVM-targeted GluCls in ubr-1.

Reduced inhibition of serotonin-stimulated pharyngeal activity by IVM in ubr-1 mutants.

(A) Left: Schematic diagram of the pharynx motor circuit composed of an excitatory cholinergic motor neuron MC, an inhibitory glutamatergic motor neuron M3 and a group of compact muscles. Both types of motor neurons could be activated by serotonin (5-HT). Right: Representative pharynx muscle Ca2+ response and kymograph evoked by 5-HT (20 mM). The dashed lines aligned with the arrows denote the initiation of perfusion. (B) Average Ca2+ traces evoked by different concentrations of 5-HT. (C) Dose-response curve of 5-HT for the pharynx muscle Ca2+ response in wild-type animals (EC50 = 20.1 mM). n ≥ 5 animals in each group. (D) Baseline-subtracted kymograph of the 5-HT-evoked pharynx muscle Ca2+ response in wild-type and ubr-1 mutants. IVM (3 μM) completely inhibited the wild-type pharynx Ca2+ response, which was significantly reduced in ubr-1 mutants. (E) Average Ca2+ traces evoked by 5-HT in different genotypes with or without IVM. (F) The dose-dependent inhibitory effect of IVM on pharynx Ca2+ was attenuated in ubr-1 mutants (blue line) compared with wild-type animals (black line) (IC50 = 0.5 μM in WT, while IC50 = 10.1 μM in ubr-1). (G) The decreased IVM inhibition of serotonin-evoked pharynx activity in ubr-1 was rescued by the restoration of UBR-1 expression (+ UBR-1), genetic removal of got-1 and eat-4, and pretreatment with ceftriaxone (+ Cef, 50 μg/mL). (H) IVM-mediated inhibition of serotonin-evoked pharynx activity in wild-type animals was blocked by glutamate feeding (WT + Glu, 20 mM). n = 10 animals in each group. (I) Quantification of the 5-HT-evoked peak amplitude of the pharynx muscle Ca2+ response in the different genotypes. Black stars: ****p < 0.0001 was used to compare intragroup differences or pharmacological groups via Student’s t test; apricot stars: ****p < 0.0001 was used to compare the differences with ubr-1 in the “+ 5-HT & IVM” group via one-way ANOVA. Error bars, SEM.

Interestingly, we observed that 5-HT induced similar Ca2+ transients in both wild-type and ubr-1 mutant worms. Furthermore, upon application of IVM (3 μM), 5-HT-elicited Ca2+ activity was profoundly suppressed in the wild-type pharynx (Figure 5D), mirroring its effect on pharyngeal pumping. However, ubr-1 mutants exhibited marked resistance to the suppressive effects of IVM, with only a modest decrease in the amplitude of Ca2+ transients (Figure 5E). Essentially, the capacity of IVM to inhibit 5-HT-driven pharynx activity was significantly compromised in ubr-1 mutants. Additionally, consistent with its behavioral impact, IVM dose-dependently suppressed 5-HT-elicited pharynx Ca2+ activity, with an estimated IVM IC50 of approximately 0.5 μM for wild-type and 10.1 μM for ubr-1 mutants (Figure 5F). These findings reveal the functional deficits of IVM-targeted GluCls in ubr-1.

In line with our earlier behavioral findings, the restoration of UBR-1 expression successfully reversed the inhibition of 5-HT-induced pharynx Ca2+ by IVM in ubr-1 mutants (Figure 5G). This corrective effect was similarly observed in ubr-1; got-1 and ubr-1; eat-4 double mutants (Figure 5G, I), reinforcing the association between decreased GluCl expression and attenuated IVM inhibition in ubr-1 mutants. These results suggest a mechanism whereby elevated glutamate levels in ubr-1 mutants may lead to glutamate-induced downregulation of GluCls, ultimately leading to IVM resistance.

To substantiate this proposed mechanism, we evaluated pharyngeal Ca2+ activity in ubr-1 mutants pretreated with Cef and in wild-type animals pretreated with glutamate. Consistent with our behavioral data, Cef treatment effectively reversed the inhibitory effect of IVM on 5-HT-evoked Ca2+ activity in the pharynx of ubr-1 mutants (Figure 5G, I), whereas glutamate exposure resulted in significant IVM resistance in wild-type animals (Figure 5H, I). This alignment between behavioral outcomes and Ca2+ activity measurements further supports the notion that glutamate accumulation is a critical factor in the downregulation of GluCls and the ensuing development of IVM resistance in ubr-1 mutants.

Discussion

In this study, we identified a crucial role of the E3 ubiquitin ligase UBR-1 in C. elegans, which contributes to IVM resistance. Given the importance of UBR-1 in regulating glutamate metabolism, our research introduces a novel mechanism for IVM resistance that could be applicable to other macrocyclic lactone (ML) anthelmintic resistance scenarios. Notably, this glutamate-based mechanism operates upstream, affecting the expression levels of GluCls, a departure from the previously documented variations in detoxification transporters that facilitate drug efflux. Through detailed investigation, we demonstrated that UBR-1 is instrumental in maintaining physiological glutamate homeostasis, thereby governing the abundance of various glutamate receptors. In ubr-1 mutants, perturbations in glutamate metabolism lead to reduced expression of primary IVM-sensitive GluCls, leading to decreased IVM-induced inhibition of serotonin (5-HT)-evoked pharynx Ca2+ activity and subsequent resistance to IVM.

A working model for UBR-1-mediated IVM resistance in C. elegans

Building upon our findings, we propose a working model that illustrates the role of the E3 ubiquitin ligase UBR-1 in mediating IVM resistance (Figure 6). In C. elegans, UBR-1 ensures functional glutamate homeostasis by inhibiting the activity of transaminase GOT-1 through a pathway that has yet to be resolved. A deficiency in UBR-1 leads to uncontrolled production of excess glutamate by GOT-1, which is then packed into synaptic vesicles for release via the vesicular glutamate transporter VGLUT/EAT-4. This excess glutamate results in the downregulation of glutamate receptors, including the IVM-targeted GluCls AVR-15, AVR-14, and GLC-1. The consequent reduction in these receptors diminishes the inhibitory effect of IVM on 5-HT-evoked pharynx Ca2+ activity, highlighting the notable IVM resistance in ubr-1 mutants. Furthermore, the exogenous application of glutamate induces IVM resistance in wild-type animals, whereas the pharmacological enhancement of glutamate transporter activity via ceftriaxone mitigates resistance in ubr-1 mutants. These observations support the conclusion that IVM resistance in ubr-1 mutants is intricately linked with aberrant glutamate metabolism.

A working model.

In wild-type animals, functional UBR-1 helps maintain balanced glutamate levels by inhibiting transaminase GOT-1 activity. In contrast, in ubr-1 loss-of-function mutants, the absence of GOT-1 inhibition leads to excessive glutamate production. This glutamate excess, which can be reduced by ceftriaxone treatment, induces compensatory downregulation of IVM-targeted GluCls. This downregulation results in a diminished inhibitory response to IVM in the pharyngeal region, ultimately causing IVM resistance in ubr-1 mutants.

Unveiling a novel mechanism of anthelmintic resistance: aberrant glutamate metabolism

Glutamate is a central metabolic hub that connects glucose and amino acid metabolism in neurons and astrocytes. Disruptions in glutamate metabolism can precipitate an overabundance of glutamate, which in turn may trigger activity-dependent modulation of glutamate receptors. This phenomenon has been observed in Drosophila, where an increase in presynaptic glutamate concentrations leads to a notable reduction in both the number of glutamate receptors and the size of the synaptic field (41). Similarly, glutamate treatments have been shown to cause a loss of postsynaptic glutamate receptors (42). In mammalian systems, prolonged exposure to glutamate has been demonstrated to specifically reduce RNA editing levels of AMPA receptors in primary cortical neurons (43). These findings suggest that a compensatory downregulation mechanism may be responsible for the observed decrease in glutamate receptors in response to excess glutamate.

In our study, we revealed that an overproduction of glutamate resulting from ubr-1 mutations leads to the downregulation of IVM-targeted GluCls, which is subsequently linked to IVM resistance in C. elegans. This contrasts with previously elucidated mechanisms of IVM resistance, such as alterations in IVM-receptor binding sites that diminish their affinity, null mutations in IVM-targeted receptors, the upregulation of cellular efflux pumps, and enhanced drug degradation (2). Our research introduces a novel metabolic mechanism involving glutamate receptors. We demonstrated that the levels of IVM-targeted receptors are subject to modulation via the glutamate metabolic pathway. This finding highlights the potential for glutamate accumulation, whether through metabolic dysregulation or exogenous administration, to confer IVM resistance, thereby offering fresh perspectives on the mechanisms behind anthelmintic resistance. Our findings therefore reveal a new UBR-1-mediated pathway via GluCls that connects glutamate metabolism with IVM resistance.

Given the high degree of conservation in glutamate pathways between C. elegans and parasitic nematodes, the implications of our findings are far-reaching. This highlights the importance of developing new pharmacological agents aimed at modulating glutamate metabolism for the control of parasitic nematodes. Moreover, altering environmental glutamate concentrations—whether through modifications in the physiological environment of nematodes or changes in the host’s diet—may increase the effectiveness of traditional macrocyclic lactone anthelmintics.

Expanding parasitic control through UBR-1 and glutamate clearance

Although it remains to be seen what percent of Ivermectin-resistant parasites in the wild have disrupted glutamate homeostasis, there is a need for genetic markers of ivermectin resistance in livestock parasites that can be used to better track resistance and to tailor drug treatment. The discovery of UBR-1 as a resistance gene in C. elegans will provide a candidate marker that can be followed up in parasitic control.

Pharmacologically, the recommendation for combination anthelmintic treatments to slow the emergence of anthelmintic resistance is well established, as they cover a similar range of parasites but act via different mechanisms (2, 44). Our research proposes an innovative approach for parasite prevention: the clearance of glutamate. The use of ceftriaxone, which has been shown to completely recover IVM sensitivity in ubr-1 mutants, presents a viable option to mitigate anthelmintic resistance prompted by excess glutamate. A combination treatment that includes ceftriaxone and IVM-like anthelmintics is advisable. Moreover, the concurrent use of ceftriaxone with other antiparasitic drugs may prove effective in treating AR parasites.

This dual-treatment strategy offers several benefits, such as the widespread availability of ceftriaxone, its established FDA approval since 1982 (45–47), the relatively low dose needed to increase IVM sensitivity, and its cost-effectiveness, particularly in the developing world.

Although the dynamics of anthelmintic resistance may vary among different hosts, our study contributes a fresh outlook by revealing a new mechanism of IVM resistance and advocating alternative preventative measures. We hope that these insights will attract the attention of global health authorities such as the World Health Organization and spark additional research in the realm of parasite control.

Materials and methods

C. elegans strains and transgenic lines

C. elegans strains were cultured at 22°C on nematode growth medium (NGM) plates seeded with E. coli OP50 as a food source (48). The wild-type Bristol N2 strain was used, and the ubr-1(hp684) strains were sourced from an EMS mutagenesis screen. Additional ubr-1 alleles were generated via the CRISPR-Cas9 system (29). Other genetic mutants were obtained from the Caenorhabditis Genetics Center (CGC). Transgenic worms were generated through microinjection via standard protocols, with target DNA plasmids at ∼50 ng/µL coinjected with the marker plasmid Pmyo-2::RFP at 5–10 ng/µL. Table S1 shows a detailed list of the strains used in this study.

Ivermectin sensitivity assay

Ivermectin and avermectin (product number: PHR1380/31372) were purchased from Sigma Aldrich (St. Louis, MO, USA) and then diluted in dimethylsulfoxide (DMSO). Doramectin (product number: D127678) was purchased from Aladdin (Shanghai, China). A stock solution of 40 mM IVM was prepared in 1 mL of DMSO. Initial ML resistance assays utilized IVM, avermectin, and doramectin at 5 ng/mL. Standard NGM plates with OP50 and 0.03% DMSO were used, with varying IVM concentrations for tests. L4 hermaphrodites were synchronized and exposed to IVM for viability assays. After 20 hours, the body length, pharyngeal pump rate, and locomotion velocity of the survivors were analyzed via a stereomicroscope (Axio Zoom V16, Zeiss). Each experiment was independently replicated at least three times.

Animal viability and motor analysis

Animal viability was assessed by calculating the relative residual ratio: the number of viable worms after 20 hours of IVM incubation divided by the initial worm count. Viability (%), body length (mm), and velocity (µm/s) were measured via a custom MATLAB script (49).

For movement analysis, IVM-exposed worms were transferred to fresh OP50 plates. For locomotion velocity analysis, a single hermaphrodite was transferred to a 60 mm imaging plate. One minute after the transfer, a three-minute video of the crawled animal was recorded on a modified stereomicroscope (Axio Zoom V16, Zeiss) with a digital camera (acA2500-60um, Basler). Postimaging analyses utilized an in-house written MATLAB script. The central line was used for tracking. Images for velocity analysis from each animal were divided into 33 body segments. The midpoint was used to calculate the body length and velocity between each frame. The pharyngeal pumping frequency was directly calculated manually via video as pumps per minute (min-1). Behavioral assays (including pumping frequency) were conducted on standard culture plates with freely moving worms.

Pharmacological pretreatment

For glutamate application, glutamate (product number: G1251, Sigma Aldrich) was dissolved in NGM at final concentrations of 0.2 mM, 1 mM, 5 mM, 10 mM, 20 mM, and 40 mM. Wild-type and ubr-1 animals were exposed to glutamate from the L1 to L4 stages and then assessed after 20 hours on glutamate and IVM plates. The same experimental procedure was used for γ-aminobutyric acid (5 mM, product number: A2129, Sigma Aldrich) and aspartate (5 mM, product number: A6202, Macklin, Shanghai, China).

In parallel, ubr-1 mutants were treated with ceftriaxone sodium (product number: TC6154, Biobomei, Hefei, China) at final concentrations of 0.05 µg/mL, 0.5 µg/mL, 5 µg/mL, 50 µg/mL, and 500 µg/mL. Viability and motor ability were evaluated after L4 worms were transferred to plates containing ceftriaxone and ivermectin (5 ng/mL) for 20 hours.

Molecular biology and RNA interference

All expression and rescue plasmids were generated via the multisite gateway system (50). Three entry clones (Slot1, Slot2, Slot3) corresponding to the promoter, target gene and fluorescent marker genes were recombined into pDEST™ R4-R3 Vector II via the LR reaction, and expression constructs were subsequently obtained. The expression patterns of ubr-1, avr-15, avr-14 and glc-1 were examined via promoter elements of 1.6 kb, 2.7 kb, 3.0 kb and 1.14 kb, respectively. DNA fragments were PCR-amplified from N2 genomic DNA.

Double-stranded RNA (dsRNA) interference technology was used to suppress gene expression in specific tissues. Essential to this process is the expression of dsRNA, which is critical for achieving targeted neuronal or tissue-specific RNAi in C. elegans (51). The plasmid for RNAi was obtained by constructing Slot2 via the BP reaction of a multisite gateway system. The exon-rich region is usually selected as the interfering region (500-700 bp), the sense strand is selected as the target gene for forward interfering sequences (RNAi, sense), and the antisense strand is selected as the target gene for reverse interfering sequences (RNAi, antisense).

Detailed plasmid and primer information is available in Tables S2-S3.

RNA sequencing analysis

For RNA-seq analysis, synchronized L4 worms were washed with bath solution at least 3 times and harvested via centrifugation. Total RNA was isolated via the freeze-thaw method, and different samples were provided to Novogene for RNA-seq analysis. Transcripts per million (TPM), a standard gene expression normalization method that adjusts the expression of a gene to the number of transcripts per million (52), was used to indicate quantitative gene expression. The TPM value takes into account the length of the gene and the sequencing depth and converts the expression of a gene to the number of transcripts per million by dividing the count value of each gene by its length and normalizing appropriately.

Fluorescence microscopy

L4 stage worms were immobilized with 2.5 mM levamisole (Sigma-Aldrich, USA) on thin layers of agarose for imaging with a laser scanning confocal microscope (Olympus FV3000, Japan) with a 60× oil objective (numerical aperture = 1.35). The intensities of AChR and GABAAR in strains EN208 kr208 [Punc-29::UNC-29::tagRFP] and KP5931 nuIs283 [Pmyo-3::UNC-49::GFP] were measured. ImageJ software was used to process the images and perform intensity analysis, and the results are reported as the normalized fluorescence intensity per individual and are shown as the mean ± SEM.

In situ Ca2+ imaging

The strain SGA616 [Pmyo-2::GCaMP6s::wCherry] was used to facilitate calcium imaging of the pharyngeal muscles. One-day-old adult hermaphrodites were glued and transferred to bath solution for imaging with a 60x water objective (Nikon, Japan; numerical aperture = 1.0). In the experiment, the epidermis behind head of the worm was gently punctured with a glass electrode (1-3 MΩ) to relieve internal pressure, which helps stabilize the calcium imaging. Serotonin (5-HT, product number: S31021, Yuanye, Shanghai, China) and ivermectin stock solutions were diluted to the desired concentrations via the following bath solution (in mM): NaCl 150; KCl 5; CaCl2 5; MgCl2 1; glucose 10; sucrose 5; and HEPES 15, pH 7.3 with NaOH, ∼330 mOsm. The control bath solution, 5-HT alone, or a combination of 5-HT and ivermectin were administered 30 seconds after initiating the recording via a gravity-driven perfusion system (INBIO MPS-3, China). Fluorescence images were acquired with an excitation wavelength LED at 470 nm and a digital sCMOS camera (Hamamatsu ORCA-Flash 4.0V2) under a wide-field microscope (Nikon LV-TV) at a rate of 10 frames per second for 2 minutes. Data were collected from HCImage (Hamamatsu) and analyzed via Image-Pro Plus 6.0 (Media Cybernetics, Inc., Rockville, MD, USA) and ImageJ (National Institutes of Health). The fluorescence intensity of the region of interest (ROI) was defined as F, and the background intensity near the ROI was defined as F0. The true muscle calcium fluorescence signal was obtained by subtracting the background signal from the ROI. ΔF/F0 = (F - F0)/F0 was plotted over time as a fluorescence variation curve.

Statistical analysis

GraphPad Prism 8 was used for statistical analyses and graphing. Dose-response residual ratio (DRR), IC50, and EC50 values were determined via a sigmoidal Dose-response model. Unpaired Student’s t test was used to analyze data between two groups, and ordinary one-way or two-way ANOVA was performed for statistical analysis when multiple groups of data were compared. The p values are indicated as follows: ns, not significant, *p < 0.05, **p < 0.01, ***p < 0.001, ****p < 0.0001. The error bars represent the SEMs.